---
title: "Thompson field day 2017"
author: "Harly Durbin"
date: 'Last updated: `r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(cowplot)
library(nlme)
library(ggjoy)
library(lubridate)
```

Need to include creation of master data frame 

```{r, eval=TRUE, echo=TRUE}
#Pull in 2016 herd report
trc_pheno <- read_excel("~/googledrive/research_extension/extension/AngusOnlineCustomPerformanceHerdReport_20170905.xlsx")

#Recode sex in order to be compatible with existing data
trc_pheno$Sex <- gsub("B", "M", trc_pheno$Sex)
trc_pheno$Sex <- gsub("C", "F", trc_pheno$Sex)

#Pull Thompson farm samples from master list, join to herd report 
trc_hair <- master %>%
  filter(Farm_ID == "TRC")
 
#Join Thompson farm herd report phenotypes to pre-existing Thompson farm hair shedding scores
trc <- full_join(trc_hair, trc_pheno, by= c("Animal_ID" = "DamTag")) %>%
  #Only keep females scored in 2016
  mutate(HairScore2016 = as.integer(HairScore2016))%>%
  mutate(HairScore2017 = as.integer(HairScore2017)) %>%
  plyr::rename(c(WNAdjWT = "Adj_wean_wt", AnmBirthDate = "Birth_date")) %>%
  mutate(Birth_date = mdy(Birth_date)) %>%
  mutate(birth_year = year(Birth_date))

#Pull out info for dams with calves in 2016, 2017
calves16 <-  trc %>%
  filter(birth_year == 2016) %>%
  select(HairScore2016, WNRatio, DamWeanAvgRatio, Reg, Adj_wean_wt, WNWT)


calves17 <- filter(trc, birth_year == 2017)

```

* Determine sample size of each score group
```{r, eval=TRUE, echo=TRUE}
trc %>% 
  group_by(HairScore2016) %>%
  tally()

trc %>% 
  group_by(HairScore2017) %>%
  tally()
```


* Plot sample size 
```{r, eval=TRUE, echo=TRUE}

score_dist17 <- ggplot(trc, aes(x = HairScore2017)) +
  geom_histogram(alpha = .4, binwidth = .5, fill = "goldenrod") +
  theme(legend.position = c(0.85, 0.85)) +
  xlab("TRC 2017 Hair Scores") +
  ylab("")

ggsave("~/googledrive/research_extension/extension/trc_score_dist.png", plot = score_dist17, device = "png")

score_dist17
```

* Plotting weaning ratio against dam hair score
```{r}
wnratio16 <- ggplot(calves16, aes(x = HairScore2016, y = DamWeanAvgRatio )) +
  geom_point() +
  geom_jitter(height = 0.3, width = 0.3) + 
  stat_summary(fun.data = mean_se, geom = "errorbar") +
  stat_summary(fun.y = mean, geom = "point", size = 3, color = "yellow")

#ggsave("~/googledrive/research_extension/extension/wnwt16.png", plot = wnratio16, device = "png")

wnratio16
```


* Plotting unadjusted weaining weight against hair score
```{r}
wnwt16 <- ggplot(calves16, aes(x = HairScore2016, y = WNWT )) +
  geom_point() +
  geom_jitter(height = 0.3, width = 0.3) + 
  stat_summary(fun.data = mean_se, geom = "errorbar") +
  stat_summary(fun.y = mean, geom = "point", size = 3, color = "yellow")

ggsave("~/googledrive/research_extension/extension/wnwt16.png", plot = wnwt16, device = "png")

wnwt16
```


* Plotting adjusted weaning weight against Hair score
```{r}
adjwnwt16 <- ggplot(calves16, aes(x = HairScore2016, y = Adj_wean_wt )) +
  geom_point() +
  geom_jitter(height = 0.3, width = 0.3) + 
  stat_summary(fun.data = mean_se, geom = "errorbar") +
  stat_summary(fun.y = mean, geom = "point", size = 3, color = "yellow")

ggsave("~/googledrive/research_extension/extension/trc_2016_calves.png", plot = prelim16, device = "png")

prelim17 <- ggplot(calves17, aes(x = HairScore2017, y = Adj_wean_wt )) +
  geom_point() +
  geom_jitter(height = 0.3, width = 0.3) + 
  stat_summary(fun.data = mean_se, geom = "errorbar") +
  stat_summary(fun.y = mean, geom = "point", size = 3, color = "yellow")
prelim17

```
 
 
 
 
 
 
```{r, eval=TRUE, echo=TRUE}


trc_lm <- lm(Adj_wean_wt ~ HairScore2016, calves16)
anova(trc_lm)
summary(trc_lm)

test <- master %>% 
  filter(!is.na(HairScore2016) & !is.na(HairScore2017))

score <- lm(HairScore2017 ~ HairScore2016, test)
anova(score)
```

* 9/11/17: lists for Jared
```{r, echo=TRUE, eval=TRUE}

one_five <- master %>%
  filter(Farm_ID == "TRC") %>%
  filter(HairScore2017 == 1 | HairScore2017 == 5) %>%
  arrange(HairScore2017) %>%
  select(-Sold2017)

write_excel_csv(one_five, "~/googledrive/research_extension/extension/170912.thompson_one_five.csv", na = "")
```

```{r, echo=TRUE, eval=TRUE, fig.width=10, fig.height=5, fig.path='~/googledrive/research_extension/extension/', dev=c('png')}


master.copy = master
master.copy <- filter(master.copy, !grepl("42", HairScore2017))
master.copy$HairScore2016 <- as.numeric(master.copy$HairScore2016)
master.copy$HairScore2017 <- as.numeric(master.copy$HairScore2017)
cor.test(master.copy$HairScore2017, master.copy$HairScore2016, use = "pairwise.complete.obs")

cor.test(trc$HairScore2017, trc$HairScore2016, use = "pairwise.complete.obs")

xtabs(~ trc$HairScore2016 + trc$HairScore2017)

trc.new <- separate(trc, Reg, into = c("One", "YearCode", "Number"), sep = c(1, 2), remove = FALSE)
trc.YearB <- filter(trc.new, YearCode %in% c("B"))
trc.YearB.HS234 <- filter(trc.YearB, HairScore2016 > 1 & HairScore2016 < 5)

lm.YearB <- lm(WNWT ~ HairScore2016, data = trc.YearB)
summary(lm.YearB)
anova(lm.HS234 <- lm(WNWT ~ HairScore2016, data = trc.YearB.HS234))
summary(lm.HS234)

9jpeg(filename = "~/googledrive/research_extension/extension/lm.wn.png")
boxplot(trc.YearB$`WN WT` ~ trc.YearB$HairScore2016,
        ylab = "Weaning Weight", xlab = "Hair Shedding Score")
#abline(a = lm.YearB$coefficients[1],
#       b = lm.YearB$coefficients[2],
#       col = "blue")
#abline(a = lm.HS234$coefficients[1],
 #      b = lm.HS234$coefficients[2],
  #     col = "red")
dev.off()

lm.YearB.adj <- lm(Adj_wean_wt ~ HairScore2016, data = trc.YearB)
summary(lm.YearB.adj)


lm.HS234.adj <- lm(Adj_wean_wt ~ HairScore2016, data = trc.YearB.HS234)
summary(lm.HS234.adj)

jpeg(filename = "~/googledrive/research_extension/extension/lm.adjwn.png")
boxplot(trc.YearB$Adj_wean_wt ~ trc.YearB$HairScore2016,
        ylab = "Adjusted Weaning Weight", xlab = "Hair Shedding Score")
#abline(a = lm.YearB.adj$coefficients[1],
#       b = lm.YearB.adj$coefficients[2],
#       col = "blue")
#abline(a = lm.HS234.adj$coefficients[1],
#       b = lm.HS234.adj$coefficients[2],
#       col = "red")
dev.off()




#ggsave("~/googledrive/research_extension/extension/lm.adjwn.png", plot = wn.adj, device = "png")
```

There is an 18 pound difference in calf weaning weights between dams with hair shedding score of 2 versus 4.
If we only analyze data from cows with hair shedding scores of 2, 3, and 4, there is a 40 pound difference.
This is due to earlier shedding cows having older and heavier calves.

When we look at adjusted weaning weights, we can remove the effect of older calves. Even on this basis, earlier shedding cows wean heavier calves. Between cows with a score of 4 and 2, the cows with a hair shedding score of 2 wean calves that are 10 pounds heavier. If we remove 1s and 5s from the analysis, the difference of weaned pounds between hair shedding scores 2 versus 4 is 26 pounds.


#PROC CORR
```{bash}
/* The CORR procedure computes Pearson correlation coefficients and the probabilities associated with these statistics. The correlation statistics include the following:
Pearson product-moment correlation: parametric measure of a linear relationship between two variables. */



/* Data from Table 5.1 in Kleinbaum test */

DATA SBP;
   INPUT PERSON SBP AGE;
   DATALINES;
 1     144 39
 2     220 47
 3     138 45
 4     145 47
 5     162 65
 6     142 46
 7     170 67
 8     124 42
 9     158 67
10     154 56
11     162 64
12     150 56
13     140 59
14     110 34
15     128 42
16     130 48
17     135 45
18     114 17
19     116 20
20     124 19
21     136 36
22     142 50
23     120 39
24     120 21
25     160 44
26     158 53
27     144 63
28     130 29
29     125 25
30     175 69
;
RUN;


/* Displays all the variables in the SAS Dataset SBP */

PROC PRINT data=SBP;
RUN;

/* Scatter plots */

proc gplot data=SBP;
plot SBP*AGE;
run;


/* Correlation */
proc corr data=SBP nomiss pearson ; /* NOMISS: Excludes observations with missing values from the analysis */
var SBP AGE;
run;

proc corr data=SBP fisher (rho0 = 0.5 alpha = 0.01) ;
/* FISHER: requests confidence limits and p-values under a specified null hypothesis, H0: rho = rho0 vs rho ? rho0, for correlation coefficients by using Fishers  transformation.  */
var SBP AGE;
run;

```

```{bash}
/* Scatter plots */

proc gplot data=thompson16;
plot HairScore2016*BirthWT;
run;


/* Correlation */
proc corr data=thompson16 nomiss pearson ; /* NOMISS: Excludes observations with missing values from the analysis */
var HairScore2016 BirthWT;
run;

proc corr data=thompson16 fisher (rho0 = 0.5 alpha = 0.01) ;
/* FISHER: requests confidence limits and p-values under a specified null hypothesis, H0: rho = rho0 vs rho ? rho0, for correlation coefficients by using Fishers  transformation.  */
var HairScore206 BirthWT;
run;

```

#Regression diagnostics
```{bash}
PROC UNIVARIATE;
VAR HairScore2016 WNAge WNAdjWT;
RUN;

PROC REG;
MODEL WNAdjWT = HairScore2016 WNAge;
PLOT rstudent.*p.;
OUTPUT out = two rstudent=jackknife_res cookd = cooksdistance h=h;
RUN;

/* rstudent = jackknife residual; cookd = cooks distance; h = leverage */

PROC UNIVARIATE data = two normal plot;
VAR jackknife_res;
PROBPLOT jackknife_res / normal;
RUN;

/* normality check for jackknife residual */

DATA outliers;
SET two;
IF (abs(jackknife_res) > tinv(0.95  8)) OR
(cooksdistance > 1) OR
(h > 0.36);
RUN;
PROC PRINT;
RUN;
```


```{bash}

```

