---
title: "AGI transfers"
author: "Harly Durbin"
date: "10/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyr)
library(stringr)
library(dplyr)
library(lubridate)
library(purrr)
library(readxl)
library(tidylog)

source(here::here("source_functions/animal_table.R"))
source(here::here("source_functions/tissue_table.R"))
```

# Setup

```{r}
animal_table <- get_animal_table("190907")

```


```{r}

transferred <-
  list.files(
    "~/Box Sync/HairShedding/ReportedData/SendToBreedAssociations",
    pattern = "\\.AN\\.xlsx|\\.AN_DNA\\.xlsx",
    full.names = TRUE
  ) %>%
  # Name the elements of the list based on a stripped down version of the filepath
  set_names(nm = (basename(.) %>%
                    tools::file_path_sans_ext())) %>%
  map(read_excel, col_types = "text", sheet = 1, trim_ws = TRUE) %>%
  # Based on the name of the file, create a column for when they were sent to Geneseek
  imap(~ mutate(.x,
                transfer_date = str_extract(.y,
                                        "^[[0-9]]+(?=\\.)"))) %>%
  # Reduce the list of dataframes to one dataframe
  reduce(bind_rows) %>%
  mutate(transfer_date = lubridate::ymd(transfer_date)) %>% 
  # Select the columns I want
  select(Lab_ID, transfer_date) %>%
  mutate(Lab_ID = as.integer(Lab_ID)) %>% 
  group_by(Lab_ID) %>% 
  filter(transfer_date == max(transfer_date))

```

```{r}
geno_check <- 
  read_rds(here::here("data/derived_data/geno_check.rds")) %>% 
  left_join(transferred)
```

# Aggregate list

* Ask about ones that aren't there but should be. Call rate/QC issues?

```{r}
geno_check %>% 
  filter(at_agi == "NO" & at_mu == "YES") %>% 
  filter(!is.na(transfer_date)) %>% 
  select(farm_id, transfer_date, call_rate, do_not_analyze, everything()) %>%
  arrange(transfer_date)
```

```{r}
geno_check %>% 
  filter(at_agi == "NO" & at_mu == "YES") %>% 
  filter(is.na(transfer_date)) %>% 
  select(farm_id, assay_mu, call_rate, do_not_analyze, everything()) %>% 
  filter(date_added > lubridate::ymd("20160901")) %>% 
  filter(is.na(do_not_analyze)) %>% 
  left_join(animal_table) %>%
  left_join(tissue %>% 
              select(Lab_ID, Source_code)) %>% 
  mutate(Ref_ID_source = "Farm_ID",
         registered = 1,
         assoc_code = "AN",
         breed_assoc = "American Angus Association") %>% 
  select(
    Lab_ID,
    Reg = registration_number,
    Barcode = Ref_ID2,
    Ref_ID,
    BC,
    Sex,
    DOB,
    Sire_Reg,
    Dam_Reg,
    registered,
    breed_assoc,
    Source_code,
    assoc_code
  ) %>% 
  writexl::write_xlsx("~/Box Sync/HairShedding/ReportedData/SendToBreedAssociations/20191009.AN.xlsx")
  
  
```

```{r}
animal_table %>% 
  filter(Lab_ID %in% c(102857, 102691))
```

# 11/7/19

"Us9001612928 FOR running water appears to be an incorrect reg # C686 is 18407898 not 18455179. So I made that correction.
 
Barcode 31815008592 for Cody Washam’s we actually on a previous project from you posted results for this sample to a tag 354 BIR 624887280 back in 2018, not reg # 17812184
 
Barcode 3116044424 for Doug Frank we actually posted results using that barcode back in 2017 on one of your projects to animal 8511 #16438724 & posted results. On this sheet you reference that sample to animal 16611219 which I think is incorrect as that animals tattoo is 0172."

```{r}
# C686 is 18407898

animal_table %>% 
  filter(Lab_ID %in% c(143307, 142869))
```

