---
title: "UMAG Animal table registration cheatsheet"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r, eval=TRUE, echo=FALSE, message=FALSE}
read_csv("~/Desktop/breeds.csv") %>% 
  filter(!is.na(Primary_American_association)) %>% 
  kableExtra::kable()
```

# Breed code (column B)

* Use the most accurate breed code BUT make sure to manually change `international_id` columns + mark as `different_international_info` if that `BC` doesn’t correspond to the breed association it’s primarily associated with 
    + I.e., Simbrah animals registered in American Simmental: mark `BC` as `SIMB` and manually change `international_id` column to reflect SIMUSA

# International IDs

Should always 19 characters long:

* Breed code (3)
* Country code (3)
* Sex code (M or F; 1)
* Registration number padded with zeros (12)

Excel formulas for creating `international_id` (column AA), `sire_international_id`, (column AF), and `dam_international_id` (column AK):

```
#Animal itself 

=Z2&Y2&G2&RIGHT("00000000000000"&X2,12)

#Sire

=AE2&AD2&"M"&RIGHT("00000000000000"&AC2,12)

#Dam

=AJ2&AI2&"F"&RIGHT("00000000000000"&AH2,12)


```

## Sire/dam interanational IDs

* If sire/dam is in closed registry, always use that primary registration
    * Closed registries:
        * Hereford
        * Angus
        * Brahman
        * Charolais (kind of… there’s an auxiliary registry for crossbreds)
        * Brangus (kind of… Ultrablacks have purebred Angus sires)
            * 
    * What about registries outside the U.S. we don’t work with, i.e., Canadian Angus bulls sire to American Simmental progeny? Use ASA registration number then?
* If sire/dam is purebred and in multiple open registries, use that breed
    * I.e., accessioning a Simmental progeny with full-blood Red Angus sire: use the Red Angus registration number in Sire_reg
* If sire/dam is crossbred and in multiple open registries, use the registration number corresponding to the breed association the accession (animal itself) is registered in
