---
title: "Thompson sample troubleshooting"
author: "Harly Durbin"
date: "2/19/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(stringr)
library(visdat)
library(reshape2)
```

#Phenotypes and Animal table

```{r}
#was trc_dr
trc_pheno <- read_excel("/Users/harlyjanedurbin/Box Sync/HairShedding/ReportedData/FilesMissingLabIDs/DataRecording_TRC_2018.xlsx", sheet = 1, trim_ws = TRUE)

#Read in UMAG database Animal table
animal_table <- Hmisc::mdb.get("/Users/harlyjanedurbin/googledrive/Samples_180604.mdb", tables = "Animal", allow = c("_"))

animal_table <- doctoR::clear_labels(animal_table) %>% 
  mutate_if(is.factor, as.character)

#Tried to do this in a pipable fashion using the naniar package but the data frame is too large
animal_table[animal_table == ""] <- NA

```

* Previously created Animal table info for my samples

```{r}
#Read in Animal table of samples in my project
trc_finished <- read_excel("~/Box Sync/HairShedding/ReportedData/FilesMissingLabIDs/DataRecording_TRC_2018.xlsx", sheet = "Animal") %>% 
  setNames(paste0(names(.), "_fin"))
  
```


##Create full pedigree
```{r}
#Create a list of the pedigree filepaths
filepaths <- list.files("~/googledrive/research_extension/hair_shed/hair_shed/hair_project/data", pattern = "TRC_Pedigree", full.names = TRUE)

#Get pedigree information including BIR numbers (to be used as Reg numbers)
#list of dataframes; a dataframe for each file
#trc_bir
trc_ped_list <- filepaths %>% 
  #Name the elements of the list based on a stripped down version of the filepath
  set_names(nm = (basename(.) %>% 
                    tools::file_path_sans_ext())) %>% 
  #Create a list of data frames using readxl
  map_df(read_excel, col_types = "text", sheet = 1, trim_ws = TRUE) %>% 
  #Rename columns in an R friendly fashion
  janitor::clean_names()
```


```{r}
trc_ped <- bind_rows(
  #Find dams
  trc_ped_list %>%
    select(dam_tag, dam_reg_no, mgs_reg_no, mgd_reg_no, mgd_tattoo) %>%
    rename(Animal_ID = dam_tag, Reg = dam_reg_no, Sire_Reg = mgs_reg_no, Dam_Reg = mgd_reg_no, Dam_ID = mgd_tattoo) %>% 
    distinct(), 
  #Find calves
  trc_ped_list %>%
    #filter(sex != "B") %>% 
    select(anm_tag, reg_no, sire_reg_no, dam_reg_no, dam_tattoo) %>%
    rename(Animal_ID = anm_tag, Reg = reg_no, Sire_Reg = sire_reg_no, Dam_Reg = dam_reg_no, Dam_ID = dam_tattoo) %>% 
    distinct()
) 

trc_ped <- trc_ped %>% 
  #Find animals listed as dams but not as an Animal_ID entry
  filter(Dam_ID %in% dplyr::setdiff(trc_ped$Dam_ID, trc_ped$Animal_ID)) %>% 
  select(Dam_Reg, Dam_ID) %>% 
  rename(Animal_ID = Dam_ID, Reg = Dam_Reg) %>% 
  bind_rows(trc_ped) %>% 
  distinct() %>% 
  setNames(paste0(names(.), "_ped"))
  

```

#Create genotype and sample table

```{r}
#Sample table information Helen provided
trc_sample_geno <- read_excel("~/Box Sync/HairShedding/ReportedData/Thompson/UMCT_samples.xlsx", sheet = 1, col_types = "text") %>%
  #Genotype check information Bob provided
  full_join(read_csv("~/Box Sync/HairShedding/ReportedData/Thompson/180711.thompson_genotype_check.csv",
                     col_types = cols(.default = "c")) %>% 
              rename(Lab_ID = lab_id) %>% 
              filter(is.na(do_not_analyze) & !is.na(assay)) %>% 
              select(Lab_ID) %>% 
              #Set distinct Lab_IDs
              distinct() %>% 
              mutate(viable_genotype = "TRUE") 
            ) %>% 
  #Animal table database information by 
  left_join(animal_table %>% 
              mutate_all(as.character) %>% 
              select(Lab_ID:Ref_ID_source3)
            )
```
            
```{r}

#Repair animal IDs
trc_sample_geno <- trc_sample_geno %>% 
  #pull out only fertility project samples (Ref_IDs that start with 'THF')
  #Create Animal_ID column by pulling ear tag from Producer_Year_EarTag formatted Ref_ID: [,3] selects 3rd item in the list
  mutate(Animal_ID = if_else(str_detect(Ref_ID, "^THF"),
                             str_split_fixed(Ref_ID, pattern = '_', n = 3)[,3],
                             Ref_ID)) %>% 
  #Extract base eartag number: look for 1 followed by a letter then pull out all following numbers
  mutate(base_num = str_extract(Animal_ID, "(?<=1[A-Z])[:digit:]+")) %>% 
  #Remove a's and b's at the end of duplicates
  mutate(base_num = str_remove(base_num, "a"),
         base_num = str_remove(base_num, "b")) %>% 
  #If base_num is less than 3 digits long, pad with zeros in front
  mutate(base_num = str_pad(base_num, 3, side = c("left"), pad = "0")) %>%
  #Extract base year of birth letter: look for multiple digits then pull out the preceeding 1 and letter
  mutate(base_yr = str_extract(Animal_ID, "1[A-Z](?=[:digit:]+)")) %>% 
  #Create "repaired ID" by pasting together base year (letter) and base eartag number
  mutate(repaired_id = str_c(base_yr, base_num)) %>% 
  select(-starts_with("base"), -BC, -NAAB,  -AIPL, -MGS, -NAAB, -duplicate, -AIPL, -Animal_ID) %>%
  #Add "Animal table" prefix for easier downstream sorting
  setNames(paste0(names(.), "_AT")) %>% 
  filter(!is.na(repaired_id_AT))
  

trc_sample_geno %>% 
  left_join(trc_ped, by = c("Ref_ID_AT" = "Animal_ID_ped")) %>% 
  write_csv("~/Desktop/trc_sample_geno.csv", na = "")
```

#Assign Lab_ID

SURPRISE we found a bunch of samples that have already been genotyped because we didn't realize Thompson has two tissue codes (THF, UMCT)


* Fertility project samples with THF source code
    + All samples relevant to my project were genotyped and are ok to use
    + Pulled 27 TSUs from box 
```{r}

umct_samples %>% 
  #pull out only fertility project samples (Ref_IDs that start with 'THF')
  filter(str_detect(Ref_ID, "^THF")) %>% 
  #Create Animal_ID column by pulling ear tag from Producer_Year_EarTag formatted Ref_ID: [,3] selects 3rd item in the list
  mutate(Animal_ID = str_split_fixed(Ref_ID, pattern = '_', n = 3)[,3]) %>% 
  #Extract base eartag number: look for 1 followed by a letter then pull out all following numbers
  mutate(base_num = str_extract(Animal_ID, "(?<=1[A-Z])[:digit:]+")) %>% 
  #If base_num is less than 3 digits long, pad with zeros in front
  mutate(base_num = str_pad(base_num, 3, side = c("left"), pad = "0")) %>%
  #Extract base year of birth letter: look for multiple digits then pull out the preceeding 1 and letter
  mutate(base_yr = str_extract(Animal_ID, "1[A-Z](?=[:digit:]+)")) %>% 
  mutate(repaired_id = str_c(base_yr, base_num)) 


animal_table %>%
  #Pull out animal table info for samples that already have Lab_IDs
  filter(Lab_ID %in% umct_samples$Lab_ID) %>% 
  #Change all columns to characters for ease of joining
  mutate_all(as.character) %>% 
  #Join to list of samples with lab IDs
  left_join(mutate_all(umct_samples, as.character)) %>% 
  select(-starts_with("base")) %>% 
  #Select only samples that are part of my project
  right_join(mutate_all(trc_finished, as.character) %>% 
               filter(Ref_ID %in% umct_samples$repaired_id), by = c("repaired_id" = "Ref_ID")) %>% 
  #Sort columns alphabetically
  select(sort(current_vars())) 
```

```{r}
x <- c("102824", "102828", "102889", "102915", "102920", "102842", "102844", "102845", "102904", "102922", "102855", "102875", "102856", "102846", "102911", "102909", "102898", "102887", "102919", "102866", "102913", "102872", "102867", "102910", "102892", "102896", "102848", "102733")

#All samples relevant to my project were genotyped and are ok to use
trc_geno_check %>% 
  right_join(animal_table %>% 
              filter(Lab_ID %in% x)) %>% 
  select(Lab_ID, assay, do_not_analyze)
```

* `Thompson Cow ID_blood_07172014.xlsx`

```{r}
#Read in sheet with IDs and extraction inforamtion
trc_14 <- read_excel("~/Box Sync/HairShedding/ReportedData/Thompson/Thompson Cow ID_blood_07172014.xlsx", sheet = 1) %>% 
  select(ID) %>% 
  #For all IDs that start with a capital letter then multiple numbers, append a "1" to the front
  mutate(ID = if_else(str_detect(ID, "(?<=[A-Z])[:digit:]+"),
                      paste0("1", ID),
                      ID)) %>% 
  rename(Ref_ID_sheet = ID) %>% 
  left_join(trc_sample_geno, by = c("Ref_ID_sheet" = "repaired_id_AT")) %>% 
  left_join(trc_finished, by = c("Ref_ID_sheet" = "Ref_ID_fin" )) %>% 
  left_join(trc_ped, by = c("Ref_ID_sheet" = "Animal_ID_ped")) %>% 
  select(order(colnames(.)))


write_csv(trc_14, "~/Desktop/2014_TRC.csv", na = "")
```


* `Thompson_Farm_Heifers_blood(JordanThomas)2011.xlsx`

```{r}
#Read in sheet with IDs and extraction inforamtion
read_excel("~/Box Sync/HairShedding/ReportedData/Thompson/Thompson_Farm_Heifers_blood(JordanThomas)2011.xlsx", sheet = 2) %>% 
  select(`Animal ID`) %>% 
  rename(Ref_ID_sheet = `Animal ID`) %>% 
  filter(!Ref_ID_sheet %in% trc_14$Ref_ID_sheet) %>% 
  left_join(trc_sample_geno, by = c("Ref_ID_sheet" = "repaired_id_AT")) %>% 
  left_join(trc_finished, by = c("Ref_ID_sheet" = "Ref_ID_fin" )) %>% 
  left_join(trc_ped, by = c("Ref_ID_sheet" = "Animal_ID_ped")) %>% 
  select(order(colnames(.))) %>%
   write_csv("~/Desktop/2011_TRC.csv", na = "")
```

* 

```{r}
read_excel("~/Box Sync/HairShedding/ReportedData/FilesMissingLabIDs/DataRecording_TRC_2018.xlsx", sheet = "trc_sample_assignment") %>% 
  select(Animal_ID, Barcode) %>% 
  #setNames(paste0(names(.), "_raw")) %>% 
  rename(repaired_id = Animal_ID) %>% 
  #Join to 2016 stuff I already went thru
  left_join(read_excel("~/Box Sync/HairShedding/ReportedData/Thompson/2016 Thompson Heifer Repro Phenotypes.xlsx", 
                       sheet = "Animal", 
                       col_types = "text") %>% 
              setNames(paste0(names(.), "_16")),
            by = c("repaired_id" = "Ref_ID_16")) %>% 
  left_join(read_excel("~/Box Sync/HairShedding/ReportedData/Thompson/Thompson Cow ID_blood_07172014.xlsx", 
                       sheet = "Animal_complete", 
                       col_types = "text") %>% 
              setNames(paste0(names(.), "_14")), 
            by = c("repaired_id" = "repaired_id_14")) %>% 
  left_join(trc_ped, 
            by = c("repaired_id" = "Animal_ID_ped")) %>% 
  left_join(trc_sample_geno, 
            by = c("repaired_id" = "repaired_id_AT")) %>% 
  select(order(colnames(.))) %>% 
  write_csv("~/Desktop/end_me.csv", na = "")
```

# 7/17/19 F heifer updates

```{r}
fhef <- 
  read_excel("~/Box Sync/HairShedding/ReportedData/2019/DataRecording_UMCT_2019.xlsx", trim_ws = TRUE) %>%
  filter(str_detect(Animal_ID, "^1F")) 


fhef %>%
  select(-contains("Registration")) %>% 
  left_join(
    read_excel(
      here::here("/data/raw_data/TRC_Pedigree.2015-2018.4June2018.xlsx"), trim_ws = TRUE
    ) %>%
      janitor::clean_names() %>%
      filter(str_detect(anm_tag, "^1F")) %>%
      select(
        Animal_ID = anm_tag,
        RegistrationNumber = reg_no,
        SireRegistration = sire_reg_no
      ),
    by = c("Animal_ID")
  ) %>% 
  select(Farm, Breed_code, RegistrationNumber, SireRegistration, everything()) %>% 
  writexl::write_xlsx("~/Desktop/fhef.xlsx")
```

