---
title: "GWAS results annotation & enrichment"
author: "Harly Durbin"
output:
  html_document:
    toc: true
    toc_depth: 2
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(readr)
library(dplyr)
library(ggplot2)
library(purrr)
library(glue)
library(magrittr)
library(stringr)
library(tidylog)
library(GALLO)
library(gprofiler2)

options(scipen = 0)
```

# Notes & questions

* SNP1101 `freq` column is allele frequency not genotype frequency: $\frac{g_2+\frac{1}{2}g_1}{g_0+g_1+g_2}$ or $\frac{2g_{2}+g_{1}}{2(g_0+g_1+g_2)}$
* GCTA COJO requirements:
    + "`A1` needs to be the effect allele with `A2` being the other allele and `freq` should be the frequency of `A1`
    + `A1` and `A2` are A/T/C/G base calls
* COJO runs very slowly with `--cojo-gc` option + it calculated a very small inflation factor, took it out

# Setup

## SNP1101 raw GWAS results

```{r, warning = FALSE, message=FALSE}
fixed9_snp1101 <-
  read_table2(here::here("data/derived_data/snp1101/fixed9/out/gwas_ssr_fixed9_bvs_p.txt"))  %>%
  janitor::clean_names() %>%
  mutate(neglog10p = -log10(p_value),
         abs_snp = abs(b_value),
         neglog10q = -log10(qvalue::qvalue(p_value)$qvalues)) %>%
  left_join(read_table2(here::here("data/derived_data/snp1101/fixed9/out/gwas_ssr_fixed9_bvs.txt"),
                        skip = 12) %>%
              select(chr = Chr, pos = Pos, fdr = FDR_GW, contains("g")))
  
```

```{r}
breeds_snp1101 <-
  c("hfd1", "an1", "bg1", "igs1") %>% 
  purrr::set_names() %>% 
  purrr::map_dfr(~ read_table2(here::here(glue("data/derived_data/snp1101/{.x}/out/gwas_ssr_{.x}_bvs_p.txt")))  %>%
                   janitor::clean_names() %>%
                   mutate(neglog10p = -log10(p_value),
                          neglog10q = -log10(qvalue::qvalue(p_value)$qvalues)) %>%
                   left_join(read_table2(here::here(glue("data/derived_data/snp1101/{.x}/out/gwas_ssr_{.x}_bvs.txt")),
                                         skip = 12) %>%
                               select(chr = Chr, pos = Pos, fdr = FDR_GW, contains("g"))),
                 .id = "dataset")
```

```{r}
full_pass <-
  fixed9_snp1101 %>% 
  filter(neglog10q > 1) %>% 
  select(CHR = chr, BP = pos)
```

```{r}
breeds_pass <-
  breeds_snp1101 %>% 
  filter(neglog10p > 5) %>% 
  select(dataset, CHR = chr, BP = pos)
```

## GFF and GTF files

```{r}
gtf <-
  import_gff_gtf(db_file = here::here("data/raw_data/Bos_taurus.ARS-UCD1.2.101.gtf"),
                 file_type = "gtf")
```

```{r}
gff <-
  import_gff_gtf(db_file = here::here("data/raw_data/Bos_taurus.ARS-UCD1.2.QTL.gff"),
                 file_type = "gff")
```

# Annotation

## Genes

```{r}
full_gene50k <-
  find_genes_qtls_around_markers(db_file = gtf,
                                 marker_file = full_pass,
                                 method = c("gene"),
                                 marker = c("snp"),
                                 interval = 50000) 
```

```{r}
full_gene10k <-
  find_genes_qtls_around_markers(db_file = gtf,
                                 marker_file = full_pass,
                                 method = c("gene"),
                                 marker = c("snp"),
                                 interval = 10000) 
```

```{r}
full_gene0 <-
  find_genes_qtls_around_markers(db_file = gtf,
                                 marker_file = full_pass,
                                 method = c("gene"),
                                 marker = c("snp"),
                                 interval = 0) 
```

```{r}
breeds_gene50k <-
  map_dfr(.x = c("an1", "bg1", "hfd1", "igs1"),
          ~ breeds_pass %>% 
            filter(dataset == .x) %>% 
            select(-dataset) %>% 
            find_genes_qtls_around_markers(db_file = gtf,
                                           marker_file = .,
                                           method = c("gene"),
                                           marker = c("snp"),
                                           interval = 50000) %>% 
            mutate(dataset = .x))
```

```{r}
breeds_gene10k <-
  map_dfr(.x = c("an1", "bg1", "hfd1", "igs1"),
          ~ breeds_pass %>% 
            filter(dataset == .x) %>% 
            select(-dataset) %>% 
            find_genes_qtls_around_markers(db_file = gtf,
                                           marker_file = .,
                                           method = c("gene"),
                                           marker = c("snp"),
                                           interval = 10000) %>% 
            mutate(dataset = .x))
```

```{r}
breeds_gene0 <-
  map_dfr(.x = c("an1", "bg1", "hfd1", "igs1"),
          ~ breeds_pass %>% 
            filter(dataset == .x) %>% 
            select(-dataset) %>% 
            find_genes_qtls_around_markers(db_file = gtf,
                                           marker_file = .,
                                           method = c("gene"),
                                           marker = c("snp"),
                                           interval = 0) %>% 
            mutate(dataset = .x))
```

```{r}
all_gene10k <-
  full_gene10k %>% 
  mutate(dataset = "full") %>% 
  bind_rows(breeds_gene10k)
```

```{r}
all_gene50k <-
  full_gene50k %>% 
  mutate(dataset = "full") %>% 
  bind_rows(breeds_gene50k)
```

```{r}
full_gene10k %>% 
  group_by(chr, gene_id, gene_name) %>% 
  tally(sort = TRUE)
```

## QTLs

```{r}
full_qtl10k <-
  find_genes_qtls_around_markers(db_file = gff,
                                 marker_file = full_pass,
                                 method = c("qtl"),
                                 marker = c("snp"),
                                 interval = 10000) 
```

```{r}
breeds_qtl10k <-
  purrr::map_dfr(.x = c("an1", "bg1", "hfd1", "igs1"),
                 ~ breeds_pass %>% 
                   filter(dataset == .x) %>% 
                   select(-dataset) %>% 
                   find_genes_qtls_around_markers(db_file = gff,
                                                  marker_file = .,
                                                  method = c("qtl"),
                                                  marker = c("snp"),
                                                  interval = 10000) %>% 
                   mutate(dataset = .x))
   
```

```{r}
breeds_gene_enrich10k[[1]]$result %>% 
  View()
```


# Calculate enrichments

## Genes

```{r}

full_gene_enrich10k <-
  full_gene10k %>% 
  mutate(id = if_else(!is.na(gene_name), gene_name, gene_id)) %>%
  distinct(id) %>% 
  pull(id) %>% 
  as.list() %>% 
  purrr::set_names() %>% 
  gost(organism = "btaurus",
       significant = TRUE)
```

```{r}

breeds_gene_enrich10k <-
  purrr::map(.x = c("an1", "bg1", "hfd1", "igs1"),
             ~ breeds_gene10k %>% 
               filter(dataset == .x) %>% 
               mutate(id = if_else(!is.na(gene_name), gene_name, gene_id)) %>%
               distinct(id) %>% 
               pull(id) %>% 
               as.list() %>% 
               purrr::set_names() %>% 
               gost(organism = "btaurus",
                    significant = TRUE))
  
```

```{r}
shared_gene_enrich_10k <-
  all_gene10k %>% 
  group_by(gene_id) %>% 
  filter(n_distinct(dataset) > 1) %>% 
  ungroup() %>% 
  mutate(id = if_else(!is.na(gene_name), gene_name, gene_id)) %>%
  distinct(id) %>% 
  pull(id) %>% 
  as.list() %>% 
  purrr::set_names() %>% 
  gost(organism = "btaurus",
       significant = TRUE)

```

## QTLs

```{r}

full_qtl_enrich10k <-
  qtl_enrich(qtl_db = gff,
             qtl_file = full_qtl10k,
             qtl_type = "Name",
             enrich_type = "genome",
             padj = "BH")
```

```{r}

breeds_qtl_enrich10k <-
  purrr::map_dfr(.x = c("an1", "bg1", "hfd1", "igs1"),
                 ~ breeds_qtl10k %>% 
                   filter(dataset == .x) %>% 
                   select(-dataset) %>% 
                   qtl_enrich(qtl_db = gff,
                              qtl_file = .,
                              qtl_type = "Name",
                              enrich_type = "genome",
                              padj = "BH") %>% 
                   mutate(dataset = .x))
  
```

# Gene ontology enrichment results

```{r}
shared_gene_enrich_10k$result
```


```{r}
gostplot(full_gene_enrich10k)
```

```{r}
publish_gosttable(full_gene_enrich10k)
```

## Testosterone?

```{r}
all_go_results %>% 
  group_by(term_name) %>% 
  filter(str_detect(term_name, "testosterone")) 
  
```

# QTL enrichment results

## Full dataset

```{r}
full_qtl_enrich10k %>% 
  View()
  filter(0.05 >= adj.pval) %>% 
  arrange(adj.pval) %>% 
  select(adj.pval, everything()) %>% 
  View()
```

## Breed-specific datasets

```{r}
breeds_qtl_enrich10k %>% 
  filter(0.05 > adj.pval) %>% 
  select(dataset, adj.pval, everything()) %>% 
  arrange(dataset, adj.pval) %>% 
  #filter(N_QTLs > 1) %>% 
  select(dataset:N_QTLs_db) %>% 
  View()
```

## White spotting?

```{r}
full_qtl10k %>% 
  filter(trait_ID == "White spotting") %>% 
  select(-start_pos, -end_pos) %>% 
  left_join(full_gene0 %>% 
              filter(CHR == 22) %>% 
              select(CHR, BP, gene_name, gene_id, start_pos, end_pos)) %>% 
  filter(between(BP, start_pos, end_pos))
  
  
  
```

```{r}
full_gene0 %>% 
  filter(CHR == 22) %>% 
  filter(between(BP, start_pos, end_pos))
```


```{r}
full_qtl10k %>% 
  filter(trait_ID == "Eye area pigmentation") %>% 
  select(-start_pos, -end_pos) %>% 
  left_join(full_gene0 %>% 
              filter(CHR == 22) %>% 
              select(CHR, BP, gene_name, gene_id, start_pos, end_pos)) 
  
  
  
```

# Chromosome 5

## Lead peak

```{r}
fixed9_snp1101 %>% 
  filter(neglog10q > 1) %>% 
  filter(chr == 5) %>% 
  filter(21000000 > pos) %>% 
  arrange(desc(neglog10q)) %>% 
  select(neglog10q, CHR = chr, BP = pos) %>% 
  left_join(full_gene10k) %>% 
  group_by(gene_name) %>% 
  tally(sort = TRUE)
```

### Lead SNP

```{r}
fixed9_snp1101 %>% 
  filter(neglog10q > 1) %>% 
  filter(chr == 5) %>% 
  filter(pos == 18767155) %>% 
  select(CHR = chr, BP = pos) %>% 
  find_genes_qtls_around_markers(db_file = gtf,
                                 marker_file = .,
                                 method = c("gene"),
                                 marker = c("snp"),
                                 interval = 415000) 
```


## Second peak

```{r}
fixed9_snp1101 %>% 
  filter(neglog10q > 1) %>% 
  filter(chr == 5) %>% 
  filter(pos > 21000000) %>% 
  arrange(desc(neglog10q)) %>% 
  select(neglog10q, CHR = chr, BP = pos) %>% 
  left_join(full_gene0) %>% 
  group_by(gene_id, gene_name) %>% 
  tally(sort = TRUE)

```

```{r}
fixed9_snp1101 %>% 
  filter(neglog10q > 1) %>% 
  filter(chr == 5) %>% 
  filter(pos > 21000000) %>% 
  arrange(desc(neglog10q)) %>% 
  select(neglog10q, CHR = chr, BP = pos) %>% 
  left_join(full_gene10k) %>% 
  group_by(gene_id, gene_name) %>% 
  tally(sort = TRUE)

```

# Chromosome 23

```{r}
all_gene10k %>% 
  filter(CHR == 23) %>% 
  select(dataset, everything(), -strand, -width) %>% 
  filter(gene_name == "PRL")
  filter(BP>= start_pos & end_pos >= BP)
```

# Chromosome 8 BG

```{r}
breeds_snp1101 %>% 
  filter(neglog10p > 5) %>% 
  filter(dataset == "bg1") %>% 
  filter(chr == 8) %>% 
  select(neglog10p, CHR = chr, BP = pos) %>% 
  left_join(breeds_gene10k)
```

# Chromosome 9 HFD

```{r}
breeds_snp1101 %>% 
  filter(neglog10p > 5) %>% 
  filter(dataset == "hfd1") %>% 
  filter(chr == 9) %>% 
  select(neglog10p, CHR = chr, BP = pos) %>% 
  left_join(breeds_gene10k)
```

# Supplementary file

```{r}
supp1 <-
  all_gene10k %>% 
  select(-CHR) %>% 
  rename(snp_pos = BP, gene_start = start_pos, gene_end = end_pos) %>% 
  left_join(fixed9_snp1101 %>% 
              mutate(dataset = "full",
                     chr = as.character(chr)) %>% 
              filter(neglog10q > 1) %>% 
              select(dataset, snp_pos = pos, chr, snp_p = p_value) %>% 
              bind_rows(breeds_snp1101 %>% 
                          mutate(chr = as.character(chr)) %>% 
                          filter(neglog10p > 5) %>% 
                          select(dataset, snp_pos = pos, chr, snp_p = p_value))) %>% 
  mutate(within_gene = case_when(snp_pos >= gene_start & gene_end >= snp_pos ~ TRUE,
                                 TRUE ~ FALSE)) %>% 
  select(dataset, chr, snp_pos, snp_p, gene_start, gene_end, gene_id, gene_name, gene_biotype, within_gene)
```


```{r}
write_csv(supp1,
          here::here("talks_articles/s1.genes10kb.csv"),
          na = "")
```

