---
title: "Gentotype dump"
author: "Harly Durbin"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(magrittr)
library(stringr)
library(tidylog)

source(here::here("source_functions/iterative_id_search.R"))
```

# Setup

```{r, warning=FALSE, message=FALSE}
animal_table <- read_rds(here::here("data/raw_data/import_join_clean/animal_table.rds"))
```

```{r, warning=FALSE, message=FALSE}
full_ped <- read_rds(here::here("data/derived_data/3gen/full_ped.rds"))
```

```{r, warning=FALSE, message=FALSE}
cleaned <- read_rds(here::here("data/derived_data/import_join_clean/cleaned.rds"))
```

```{r, warning=FALSE, message=FALSE}
sample_table <- 
  read_csv(here::here("data/raw_data/import_join_clean/200820_sample_sheet.csv"),
           trim_ws = TRUE,
           guess_max = 100000)
```

# Generate `geno_dump` list of animals with Lab IDs

```{r, warning=FALSE, message=FALSE}
geno_dump <-
  cleaned %>% 
  distinct(Lab_ID) %>% 
  bind_rows(full_ped %>% 
              distinct(Lab_ID)) %>% 
  distinct() %>% 
  filter(!is.na(Lab_ID))
```

```{r, warning=FALSE, message=FALSE}
geno_dump %<>% 
  # Add in duplicate lab IDs
  bind_rows(geno_dump %>% 
              left_join(animal_table %>% 
                          select(duplicate, Lab_ID)) %>% 
              select(Lab_ID = duplicate) %>% 
              filter(!is.na(Lab_ID)) %>% 
              filter(Lab_ID != 0))
```

## Export `geno_dump` list

```{r, warning=FALSE, message=FALSE, eval = FALSE}
geno_dump %>% 
  write_tsv(here::here("data/derived_data/geno_dump/200923.geno_dump.txt"), col_names = FALSE)
```

# Missing genotypes

## How many animals have sample IDs and should have acceptable genotypes?

```{r, warning=FALSE, message=FALSE}
geno_dump %>% 
  left_join(sample_table %>% 
              select(Lab_ID = lab_id, sample_id, call_rate, do_not_analyze)) %>% 
  # Has sample ID
  filter(!is.na(sample_id)) %>% 
  # Not marked as do_not_analyze
  filter(is.na(do_not_analyze)) %>% 
  filter(!is.na(call_rate))
```

## Sample IDs actually dumped 

* ~~2,000 samples missing, compare ID files with list of what I expected~~
    + Ended up being issue with how Bob stored BOVGV2 genotypes

# Post-genotype dump tally

```{r, warning=FALSE, message=FALSE}
post_dump <-
  list.files(here::here("data/derived_data/geno_dump"), pattern = "ID", full.names = TRUE) %>% 
  set_names(nm = (basename(.) %>%
                    tools::file_path_sans_ext())) %>%
  purrr::map_dfr(~ read_table2(.x, col_names = c("rownum", "international_id")), .id = "file") %>% 
  distinct(international_id) %>% 
  left_join(sample_table %>% 
              select(Lab_ID = lab_id, international_id))
```

## Variation in density

```{r, warning=FALSE, message=FALSE}

list.files(here::here("data/derived_data/geno_dump"), pattern = "ID") %>% 
  set_names(nm = (basename(.) %>%
                    tools::file_path_sans_ext())) %>% 
  as_tibble() %>% 
  mutate(density = str_extract(value, "^[[:digit:]]+(?=\\.)"),
         density = as.numeric(density)) %>% 
  arrange(density)
```

```{r}
list.files(here::here("data/derived_data/geno_dump"), pattern = "ID", full.names = TRUE) %>% 
  set_names(nm = (basename(.) %>%
                    tools::file_path_sans_ext())) %>%
  purrr::map_dfr(~ read_table2(.x, col_names = c("rownum", "international_id")), .id = "file") %>% 
  mutate(density = str_extract(file, "^[[:digit:]]+(?=\\.)"),
         density = as.numeric(density)) %>% 
  group_by(density) %>% 
  tally(sort = TRUE)
```


# Post-imputation tally

```{r}

read_table2(here::here("data/derived_data/geno_dump/200924_HairShed.850K.fam"), col_names = FALSE) %>% 
  select(international_id = X1) %>% 
  distinct()
```

# Commentary

```{r, message=TRUE, warning=FALSE}
post_dump %>% 
  left_join(cleaned %>% 
              distinct(Lab_ID, farm_id, animal_id, temp_id)) %>% 
  select(-Lab_ID) %>% 
  mutate(in_project = if_else(!is.na(farm_id), "yes", NA_character_)) %>% 
  group_by(international_id) %>% 
  fill(in_project, .direction = "downup") %>% 
  ungroup() %>% 
  distinct(international_id, in_project) %>% 
  group_by(in_project) %>% 
  tally()
```

Genotype data was available for 10,709 phenotyped animals and an additional 1,089 relatives. 

---