---
title: "April 2020 e-mail breeding values"
author: "Harly Durbin"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyr)
library(stringr)
library(dplyr)
library(glue)
library(readxl)
library(readxl)
library(magrittr)
library(purrr)
library(tidylog)

source(here::here("source_functions/cg_tallies.R"))
source(here::here("source_functions/coalesce_join.R"))
source(here::here("source_functions/three_gen.R"))

```

# Notes & questions

# Setup 

```{r}
animal_table <- read_rds(here::here("data/derived_data/animal_table.rds"))
```

```{r}
cleaned <- read_rds(here::here("data/derived_data/cleaned.rds"))
```

```{r}
full_ped <- read_rds(here::here("data/derived_data/full_ped.rds"))
```

```{r}
old_samp <- 
  read_delim(here::here("data/raw_data/190205_sample_sheet.csv"),
             delim = c(","),
             col_names = TRUE,
             guess_max = 10000) %>% 
  rename(Lab_ID = lab_id,
         Reg = reg)
```

# List of Lab IDs for Bob

```{r, eval = FALSE}

progeny <- 
  cleaned %>% 
  distinct(Lab_ID) %>% 
  left_join(animal_table) %>% 
  select(Lab_ID, Reg, lab_id_sire, Sire_Reg, lab_id_dam, Dam_Reg) 

parents <- 
  progeny %>% 
  select(Lab_ID = lab_id_sire) %>% 
  bind_rows(
    progeny %>% 
      select(Lab_ID = lab_id_dam)
  ) %>% 
  filter(!is.na(Lab_ID)) %>% 
  distinct() %>% 
  left_join(animal_table) %>% 
  select(Lab_ID, Reg, lab_id_sire, Sire_Reg, lab_id_dam, Dam_Reg) 


grandparents <-
  parents %>% 
  select(Lab_ID = lab_id_sire) %>% 
  bind_rows(
    parents %>% 
      select(Lab_ID = lab_id_dam)
  ) %>% 
  filter(!is.na(Lab_ID)) %>% 
  distinct() %>% 
  left_join(animal_table) %>% 
  select(Lab_ID, Reg, lab_id_sire, Sire_Reg, lab_id_dam, Dam_Reg) 
```

```{r, eval = FALSE}
  bind_rows(progeny, parents, grandparents) %>% 
  distinct(Lab_ID) %>% 
  filter(!is.na(Lab_ID)) %>% 
  arrange(Lab_ID) %>% 
  write_tsv(here::here("data/derived_data/200402.hairshed_dump.txt"), col_names = FALSE)
```

# Contemporary grouping

```{r}
cleaned %>% 
  group_by(farm_id) %>% 
  tally(sort = TRUE)
```

First, remove males

```{r}
update_dat <-
  cleaned %>%
  # Females only
  filter(sex == "F") 
```


## Calving season

1. Figure out some with Angus calving season data
2. Then, impute using most recent calving date where possible
3. Then, use birth date

```{r}
update_dat %<>%
  left_join(read_excel(
    here::here("data/derived_data/season_search.xlsx"),
    trim_ws = TRUE,
    skip = 1,
    col_names = c("farm_id", "registration_number", "year", "cs2")
  )) %>% 
  mutate(calving_season = if_else(!is.na(cs2), cs2, calving_season)) %>% 
  # If calving season still missing, impute using most recent
  group_by(farm_id, temp_id) %>% 
  arrange(date_score_recorded) %>% 
  fill(calving_season, .direction = "downup") %>% 
  ungroup() %>% 
  # If calving season still missing, impute using DOB
  mutate(
    calving_season = 
      case_when(
          is.na(calving_season) & between(lubridate::month(dob), left = 1, right = 6) ~ "SPRING",
          is.na(calving_season) & between(lubridate::month(dob), left = 7, right = 12) ~ "FALL",
        TRUE ~ calving_season
      )) %>% 
  select(-cs2)
  
```

```{r}
update_dat %>% 
  summarise(missing = sum(is.na(calving_season)),
            found = sum(!is.na(calving_season)))
```

```{r}
update_dat %>% 
  filter(is.na(calving_season)) %>% 
  group_by(farm_id) %>% 
  tally()
```



## Age group

* Need to add Angus pedigree age updates

```{r}
update_dat %<>%
  mutate(age_group =
           case_when(
             # Heifers and yearlings
             age == 1 ~ "yearling",
             age == 2 ~ "fch",
             age == 3 ~ "three",
             age >= 4 ~ "mature"
           ))
```

```{r}
update_dat %>% 
  filter(is.na(age_group)) %>% 
  group_by(farm_id) %>% 
  tally(sort = TRUE)

```

## Score group

* Establish score groups

```{r, eval=FALSE}
update_dat %>% 
  filter(!farm_id %in% c("SAV", "BAT")) %>% 
  group_by(farm_id, year) %>% 
  filter(n_distinct(date_score_recorded) > 1) %>% 
  ungroup() %>% 
  group_by(farm_id, date_score_recorded) %>% 
  tally() %>% 
  ungroup() %>% 
  left_join(
    read_excel(here::here("data/derived_data/score_windows2.xlsx")) %>% 
      select(farm_id, date_score_recorded, score_group) %>% 
      mutate(date_score_recorded = lubridate::ymd(date_score_recorded))
  ) %>% 
  mutate(year = lubridate::year(date_score_recorded)) %>% 
  select(farm_id, year, everything()) %>% 
  writexl::write_xlsx(here::here("data/derived_data/score_groups.xlsx"))
```

```{r, eval=FALSE}
update_dat %>% 
  filter(farm_id %in% c("SAV", "BAT")) %>% 
  group_by(farm_id, year) %>% 
  filter(n_distinct(date_score_recorded) > 1) %>% 
  ungroup() %>% 
  group_by(farm_id, date_score_recorded) %>% 
  tally() %>% 
  ungroup() %>% 
  mutate(year = lubridate::year(date_score_recorded)) %>% 
  select(farm_id, year, everything()) %>% 
  writexl::write_xlsx(here::here("data/derived_data/ua_score_groups.xlsx"))
```

* Join to data

```{r}
update_dat %<>%
  left_join(
    bind_rows(read_excel(
      here::here("data/derived_data/ua_score_groups.xlsx")
    ),
    read_excel(
      here::here("data/derived_data/score_groups.xlsx")
    )) %>%
      select(farm_id, date_score_recorded, score_group) %>%
      mutate(date_score_recorded = lubridate::ymd(date_score_recorded))
  ) %>% 
  mutate(score_group = tidyr::replace_na(score_group, 1))
```

## Final contemporary groupings

```{r}
update_dat %>%
  mutate(
    cg = glue::glue(
      "{farm_id}{year}{calving_season}{age_group}{score_group}{toxic_fescue}"
    ),
    cg_num = as.integer(factor(cg))
  ) %>%
  cg_tallies()
```


```{r}

update_dat %<>%
  mutate(
    cg = glue::glue(
      "{farm_id}{year}{calving_season}{age_group}{score_group}{toxic_fescue}"
    ),
    cg_num = as.integer(factor(cg))
  ) %>%
  group_by(cg) %>%
  # At least 5 animals per CG
  filter(n() >= 5) %>%
  # Remove CGs with no variation
  filter(var(hair_score) != 0) %>%
  ungroup() %>% 
  left_join(full_ped %>%
              distinct(Lab_ID, farm_id, animal_id, temp_id, full_reg, sire_reg, dam_reg)) %>%
  # 200417 fit breed as random effect
  mutate(
    breed_effect =
      str_extract(full_reg, "^SIM|^HER|^RAN|^BSH|^CIA|^RDP|^CHA|^BGR|^AAA|^AMGV|^GVH|^HFD|^BIR|^AAN|^AMAR|^AMXX"),
    breed_effect =
      case_when(
        breed_effect %in% c("AAA", "AAN", "AN", "BIR") ~ "AN",
        breed_effect %in% c("ANR", "AMAR", "RAN") ~ "ANR",
        breed_effect %in% c("HER", "HFD") ~ "HFD",
        breed_effect %in% c("AMXX", "AMGV", "GVH", "GEL") ~ "GEL",
        breed_effect %in% c("BGR", "BG") ~ "BG",
        breed_effect %in% c("MAAN", "RDP") ~ "MAAN",
        breed_effect %in% c("SH", "BSH") ~ "SH",
        breed_effect %in% c("CHIA", "CIA") ~ "CHIA",
        is.na(breed_effect) & !is.na(breed_code) ~ breed_code,
        TRUE ~ breed_effect
      )
  ) #%>% 
  #group_by(breed_effect) %>% 
  #filter(n() >= 50) %>% 
  #ungroup()


```

# Matching pedigree IDs to genotype IDs 

* Send Brangus duplicates to Bob & Helen

```{r, eval=FALSE}
read_table(here::here("data/raw_data/190215_hair_shedding.sample"),
             col_names = "international_id") %>% 
  mutate(international_id = str_remove(international_id, "1_")) %>% 
  left_join(old_samp %>% 
              select(international_id, Lab_ID) %>% 
              distinct(international_id, Lab_ID)) %>% 
  group_by(international_id) %>% 
  #NOF Gelbvieh duplicates strike again
  filter(Lab_ID == max(Lab_ID)) %>% 
  ungroup() %>% 
  left_join(full_ped %>% 
              select(-sire_reg, -dam_reg)) %>% 
  filter(is.na(full_reg)) %>% 
  select(Lab_ID) %>% 
  left_join(animal_table) %>% 
  filter(BC == "BG") %>% 
  select(Lab_ID1 = Lab_ID, registration_number = Reg) %>% 
  left_join(cleaned %>% 
              distinct(registration_number, Lab_ID)) %>% 
  pivot_longer(cols = c("Lab_ID1", "Lab_ID")) %>% 
  select(registration_number, Lab_ID = value) %>% 
  left_join(animal_table) %>% 
  writexl::write_xlsx("~/Desktop/brangus_dup.xlsx")
```

```{r}
geno_key <-
  read_table(here::here("data/raw_data/190215_hair_shedding.sample"),
             col_names = "international_id") %>% 
  mutate(international_id = str_remove(international_id, "1_")) %>% 
  left_join(old_samp %>% 
              select(international_id, Lab_ID) %>% 
              distinct(international_id, Lab_ID)) %>% 
  group_by(international_id) %>% 
  #NOF Gelbvieh duplicates strike again
  filter(Lab_ID == max(Lab_ID)) %>% 
  ungroup() %>% 
  left_join(animal_table %>% 
              select(Lab_ID, Reg)) %>% 
  left_join(full_ped %>% 
              select(-sire_reg, -dam_reg)) %>% 
  # Coalesce in Brangus duplicates
  coalesce_join(full_ped %>% 
                  filter(str_detect(full_reg, "^BGR")) %>% 
                  mutate(Reg = str_remove(full_reg, "BGR")) %>% 
                  select(Reg, full_reg, farm_id, animal_id, temp_id, Lab_ID2 = Lab_ID),
                join = dplyr::left_join,
                by = c("Reg")) %>% 
  mutate(Lab_ID = if_else(!is.na(Lab_ID2), Lab_ID2, Lab_ID)) %>% 
  select(-Reg, -Lab_ID2)
```

## Sample ID file 

```{r}
read_table(here::here("data/raw_data/190215_hair_shedding.sample"),
             col_names = "international_id") %>% 
  mutate(international_id = str_remove(international_id, "1_")) %>% 
  left_join(geno_key %>% 
              select(international_id, full_reg)) %>% 
  mutate(full_reg = if_else(is.na(full_reg), international_id, full_reg)) %>% 
  select(full_reg) %>% 
  write_tsv(here::here("data/raw_data/190215_hair_shedding.geno_key"), col_names = FALSE)
```



# Match pedigree IDs to cleaned IDs, write out data


```{r}
update_dat %>% 
  select(full_reg, everything()) %>% 
  select(full_reg, cg_num, hair_score) %>% 
  write_delim(here::here(glue::glue("data/derived_data/update_email2020/no_breed/data.txt")),
                         delim = " ",
                         col_names = FALSE)

```

```{r}
update_dat %>% 
  left_join(full_ped %>% 
              distinct(Lab_ID, farm_id, animal_id, temp_id, full_reg)) %>% 
  select(full_reg, everything()) %>% 
  left_join(geno_key %>% 
              mutate(genotyped = TRUE)) %>% 
  select(-international_id) %>% 
  write_rds(here::here("data/derived_data/update_dat.rds"))
```


# Make pedigree

```{r}

update_dat %>% 
  left_join(full_ped %>% 
              distinct(Lab_ID, farm_id, animal_id, temp_id, full_reg, sire_reg, dam_reg)) %>% 
  mutate(full_reg = if_else(is.na(full_reg), as.character(glue("{farm_id}{animal_id}{temp_id}")), full_reg)) %>% 
  select(full_reg, sire_reg, dam_reg) %>% 
  three_gen(full_ped = full_ped) %>% 
  write_delim(here::here(glue::glue("data/derived_data/update_email2020/no_breed/ped.txt")),
                         delim = " ",
                         col_names = FALSE)
```

# Map file

* Pull out chromosome and position from mgf file
    + `zcat data/raw_data/190215_hair_shedding.hard.mgf.gz | cut -d"," -f1 > data/raw_data/190215_hair_shedding.chr_pos.txt`

```{r}
read_table2(here::here("data/raw_data/190215_hair_shedding.chr_pos.txt"), col_names = FALSE, col_types = "c") %>% 
  separate(1, into = c("chr", "pos"), sep = ":") %>% 
  mutate(snp_number = row_number()) %>% 
  select(snp_number, chr, pos) %>% 
  write_tsv(here::here("data/raw_data/190215_hair_shedding.chrinfo.txt"), col_names = FALSE)
```

