---
title: "Formatting data for `sommer` + intro"
author: "Harly Durbin"
date: "11/27/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(sommer)
library(tidyverse)
```

# Create phenotype dataframe

## Joining sample data to phenotype data/metadata

```{r}


#Read in PLINK fam file: important for sample order
fam <-
  read_table2("gemma_in/180506.f250.fam", col_names = FALSE) %>%
  dplyr::select(X2, X3) %>%
  #Trim white space
  mutate(X2 = str_trim(X2, side = c("both")), X3 = str_trim(X3, side = c("both"))) %>%
  #If international id == sire international id, change sire international id to 0
  mutate(X3 = if_else(X3 == X2, "0", X3)) %>%
  #Left join dataframe of phenotypes (gwas_master) to PLINK fam file (resulting dataframe is in order of PLINK fam file)
  left_join(
    gwas_master %>%
      dplyr::select(
        international_id,
        Farm_ID,
        CalvingSeason2016,
        CalvingSeason2017,
        Sex,
        Age2016,
        Age2017,
        DateScoreRecorded2016,
        DateScoreRecorded2017,
        DateDeviation2016,
        DateDeviation2017,
        sixty_avg_16,
        sixty_avg_17,
        HairScore2016,
        HairScore2017,
        LAT
      ),
    by = c("X2" = "international_id")
  ) %>%
  distinct() %>%
  #Change date deviation from date to integer
  mutate(
    DateDeviation2016 = as.integer(DateDeviation2016),
    DateDeviation2017 = as.integer(DateDeviation2017)
  ) %>%
  #Make a column of row numbers for filtering
  mutate(numb = row_number()) %>%
  #Remove OHG duplicate (scored twice in 2017)
  filter(numb != 4623) %>%
  #Remove from 2017 analysis (scored twice, change its 2017 hair score to NA)
  mutate(HairScore2017 = replace(HairScore2017, numb == 4622, NA))
```

## Impute missing phenotype values + removing duplicates

```{r}
#How many times does each international ID occur?
n_occur_reg <- as.data.frame(table(fam$X2))
#Pull out international IDs where # occurences > 1, make data frame of row numbers to toss
dup_reg <-
  as_data_frame(fam[fam$X2 %in% n_occur_reg$Var1[n_occur_reg$Freq > 1], ]) %>%
  filter(is.na(HairScore2016))


#Can't have missing phenotype values in order for observation to not be dropped by sommer
#mean 2016 date deviation is 27
mean_dev <- fam %>%
  filter(!is.na(DateDeviation2016)) %>%
  summarise(mean(DateDeviation2016)) %>%
  as.integer()


#mean 2016 age is 3
mean_age <- fam %>%
  filter(!is.na(Age2016)) %>%
  summarise(mean(Age2016)) %>%
  as.integer()

#mean 2016 60d avg is 16.275
mean_sixty_avg <- fam %>%
  filter(!is.na(sixty_avg_16)) %>%
  summarise(mean(sixty_avg_16)) %>%
  as.numeric()

```


```{r, eval=FALSE, echo=TRUE}

cov <- fam %>%
  #Remove if in dup_reg
  filter(!numb %in% dup_reg$numb) %>%
  #Remove row number column
  dplyr::select(-numb) %>%
  #Can't have NAs during contemporary group creation; change relevant NAs to a U
  mutate(Sex = replace(Sex, is.na(Sex), "U")) %>%
  #If any males have a calving season assigned, remove it
  mutate(
    CalvingSeason2016 = replace(CalvingSeason2016, Sex == "M", "U"),
    CalvingSeason2017 = replace(CalvingSeason2017, Sex == "M", "U")
  ) %>%
  #Change NA calving season for females and unknown sex to U
  mutate(
    CalvingSeason2016 = replace(CalvingSeason2016, is.na(CalvingSeason2016), "U"),
    CalvingSeason2017 = replace(CalvingSeason2017, is.na(CalvingSeason2017), "U")
  ) %>%
  #Change
  #Create contemporary groupings
  mutate(sex_group_16 = as.factor(str_c(Sex, CalvingSeason2016)),
         sex_group_17 = as.factor(str_c(Sex, CalvingSeason2017))) %>%
  #Change farm to a factor
  mutate(Farm_ID = as.factor(Farm_ID)) %>%
  #Change hair score to numeric
  mutate_at(vars(starts_with("Hair")), funs(as.numeric(.))) %>%
  #mean impute DateDeviation, Age, 60d avg for 2016
  mutate(
    DateDeviation2016 = replace(
      DateDeviation2016,
      is.na(DateDeviation2016) & !is.na(HairScore2016),
      mean_dev_16
    ),
    Age2016 = replace(Age2016, is.na(Age2016) &
                        !is.na(HairScore2016), mean_age_16),
    sixty_avg_16 = replace(
      sixty_avg_16,
      is.na(sixty_avg_16) & !is.na(HairScore2016),
      mean_sixty_16
    )
  ) %>%
  #mean impute DateDeviation, Age, 60d avg for 2017
  mutate(
    DateDeviation2017 = replace(
      DateDeviation2017,
      is.na(DateDeviation2017) & !is.na(HairScore2017),
      mean_dev_17
    ),
    Age2017 = replace(Age2017, is.na(Age2017) &
                        !is.na(HairScore2017), mean_age_17),
    sixty_avg_17 = replace(
      sixty_avg_17,
      is.na(sixty_avg_17) & !is.na(HairScore2017),
      mean_sixty_17
    )
  ) %>%
  mutate(
    international_id = as.factor(international_id),
    Farm_ID = as.factor(Farm_ID),
    CalvingSeason2016 = as.factor(CalvingSeason2016),
    CalvingSeason2017 = as.factor(CalvingSeason2017),
    Sex = as.factor(Sex)
  )

#Assign row names as international id (for downstream sommer analyses)
row.names(cov) <- cov$international_id

#Remove animals not in Troy's imputation run
f250 <- read_table2("pre_adjust/227234.180430.6615.A.ID", col_names = FALSE)

cov_uni <- cov %>%
  filter(international_id %in% f250$X2)

row.names(cov_uni) <- cov_uni$international_id

saveRDS(cov_uni, "peqg/phenotypes_uni.RDS")

```

## "Long" format

```{r}

#Create long format dataframe with one line for each year's observation (i.e., animals with 3 years of scores will have 3 rows in the dataframe)
#This could have been done a lot easier with dplyr::gather but didn't realize that at the time
cov_multi <- bind_rows(cov_uni %>%
                         filter(!is.na(HairScore2016)) %>% 
                         rename(HairScore = HairScore2016, DateDeviation = DateDeviation2016, CalvingSeason = CalvingSeason2016, Age = Age2016, sixty_avg = sixty_avg_16, sex_group = sex_group_16, ToxicFescue = ToxicFescue2016) %>% 
                         mutate(yr = 2016) %>% 
                         dplyr::select(international_id, Farm_ID, CalvingSeason, Sex, Age, DateDeviation, sixty_avg, norm_30y, HairScore, LAT, yr, sex_group),
                       cov_uni %>%
                         filter(!is.na(HairScore2017)) %>%
                         rename(HairScore = HairScore2017, DateDeviation = DateDeviation2017, CalvingSeason = CalvingSeason2017, Age = Age2017, sixty_avg = sixty_avg_17, sex_group = sex_group_17, ToxicFescue = ToxicFescue2017) %>%
                         mutate(yr = 2017) %>% 
                         dplyr::select(international_id, Farm_ID, CalvingSeason, Sex, Age, DateDeviation, sixty_avg, norm_30y, HairScore, LAT, yr, sex_group)) %>%
  mutate(con_group = str_c(Farm_ID, CalvingSeason, yr)) %>% 
  #could have been done easier with dplyr::mutate_at()
  mutate(international_id = as.factor(international_id), yr = as.factor(yr), Farm_ID = as.factor(Farm_ID), CalvingSeason = as.factor(CalvingSeason), Sex = as.factor(Sex), con_group = as.factor(con_group))

saveRDS(cov_multi, "peqg/phenotypes_multi.RDS")
```

# GRM

```{r, eval=TRUE, echo=FALSE}

grm <- read_table2("pre_adjust/180502.imputed.grm.sXX.txt", col_names = FALSE) %>%
  #For some reason, ended up with an extra column of just NAs, remove it
  dplyr::select(-X6453) %>% 
  drop_na() 

#Since GRM should be in same order as above .fam file and therefore cov_uni dataframe, name GRM rows and columns based on order of cov_uni
row.names(grm) <- cov_uni$international_id
colnames(grm) <- cov_uni$international_id
   
saveRDS(grm, file = "180507.grm_imputed.RDS")


#Drop rows and columns where corresponding to rows in cov_uni where HairScore2016 is NA
   
grm <- base::as.matrix(grm)

```

# Identity matrix 

```{r}
#Diagonal matrix same dimensions as GRM
pe <- base::diag(x = 1, nrow = 6452, ncol = 6452)

row.names(pe) <- row.names(grm)
colnames(pe) <- colnames(grm)
```


# `sommer`

## Tool notes

* `mmer2` is formula based
    + To specify fixed effects, `1+[fixed effects]` after phenotype specification
        - Only specify random effects inside `g()` if it has a corresponding covariance matrix (i.e., only use for animal effect/GRM)
        - Specify : `G=(list(X1=grm))`
* Repeatability/permanent environmental troubleshooting
    + ~~Add repeatability component with `~g(X1)` (i.e., random effect of individual)~~
    + Specifying a clone ID column (i.e., `g(international_id.pe)`) in random effects and an identity matrix in G does not work, also tried variations on G list formatting
    + Possible solution: using `at()`? I.e., `at(yr,c("2016", "2017"):ID)`
        - Or making my own repeatability matrix, whatever that entails

## Create toy dataset for testing

```{r}

toy_data <- cov_uni %>% 
  dplyr::filter(!is.na(HairScore2016) & !is.na(HairScore2017)) %>% 
  sample_n(200, replace = FALSE)

#filter(cov_uni, CalvingSeason2016 != CalvingSeason2017)

row.names(toy_data) <- toy_data$international_id
```

## Solve MME using `mmer2()`

```{r}
#Here, use DateDeviation, CalvingSeason, Sex, Age, and Farm_ID as fixed effects
#Each of these is a column in toy_data
#Make sure factors are coded correctely in the source data frame
pre_adjust_toy <- sommer::mmer2(HairScore~1 + DateDeviation + CalvingSeason + Sex + Age + Farm_ID,
#specify individual as a random effect
#international_id is a column in toy_data                  
              random = ~g(international_id) + g(international_id.pe),
              rcov = ~units,
#Specify the dataframe where data is coming from
              data = toy_data,
#Include the GRM (generated by GEMMA) as a random effect
#grm is a matrix object, this isn't coming from toy_data
              G = list(international_id = grm, international_id.pe = pe)
)

```

## Extract variance components/heritabilities 

```{r}
#Pull out the variance components and save it as an object
suma <- summary(pre_adjust_toy)$var.comp.table


#calculate heritability from the object you just save
sum(suma[1,1]/sum(suma[,1]))
```



