---
title: "simmental_ww_hs"
author: "Harly Durbin"
date: "7/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(magrittr)
library(tidyr)
library(readr)
library(purrr)
library(tibble)
library(forcats)
library(readxl)
library(stringr)
library(viridis)
library(cowplot)
#library(lubridate)
library(maps)

master <- readRDS("data/master.RDS")
```

# Questions

* Should I be making better use of contemporary grouping information?
* Fall calving?? Exclude those that weaned calves before they were scored?
* Is there a statistically kosher way to make use of the historical data on years' past calves to increase power?
    + Could use historical data if it was compared to breeding value
* A way to combine 2016 and 2017 observations?

#Weaning weight 

```{r}
sim_dam_16 <- read_excel("data/180706.sim.Dam_Progeny_Report.xlsx") %>% 
  janitor::clean_names() %>% 
  #Change DOB and wean date to dates
  mutate_at(c("calf_dob", "wean_date"), as.Date) %>% 
  mutate(wean_yr = str_extract(wean_date, "^[:digit:]{4}")) %>% 
  #Keep only 2017 observations
  filter(wean_yr == 2016) %>% 
  #Join progeny report to contemporary grouping data
  left_join(read_excel("data/180706.sim.Anm_CG_Data.xlsx") %>%
              janitor::clean_names() %>% 
              rename(calf_sex = sex, 
                     calf_asa_number = anm_nbr) %>%
              #Change sex codes
              mutate(calf_sex = replace(calf_sex, calf_sex == "C", "F"),
                     calf_sex = replace(calf_sex, calf_sex == "B" | calf_sex == "S", "M")) %>% 
              select(calf_sex, calf_asa_number)) %>% 
  #Join to phenotype data
  left_join(master %>% 
              select(Farm_ID, 
                     Reg, 
                     Color,
                     ends_with("2016")), 
            by = c("dam_asa_number" = "Reg")) %>% 
  filter(!is.na(CalvingSeason2016),
         !is.na(HairScore2016), 
         !is.na(calf_sex)) %>% 
  mutate(year = "2016") %>% 
  rename(Age = Age2016, CalvingSeason = CalvingSeason2016, ToxicFescue = ToxicFescue2016, HairScore = HairScore2016) %>% 
  select(-Shearing2016, - Comment2016)
```


```{r, warning=FALSE}

sim_dam_17 <- read_excel("data/180706.sim.Dam_Progeny_Report.xlsx") %>% 
  janitor::clean_names() %>% 
  #Change DOB and wean date to dates
  mutate_at(c("calf_dob", "wean_date"), as.Date) %>% 
  mutate(wean_yr = str_extract(wean_date, "^[:digit:]{4}")) %>% 
  #Keep only 2017 observations
  filter(wean_yr == 2017) %>% 
  #Join progeny report to contemporary grouping data
  left_join(read_excel("data/180706.sim.Anm_CG_Data.xlsx") %>%
              janitor::clean_names() %>% 
              rename(calf_sex = sex, 
                     calf_asa_number = anm_nbr) %>%
              #Change sex codes
              mutate(calf_sex = replace(calf_sex, calf_sex == "C", "F"),
                     calf_sex = replace(calf_sex, calf_sex == "B" | calf_sex == "S", "M")) %>% 
              select(calf_sex, calf_asa_number)) %>% 
  #Join to phenotype data
  left_join(master %>% 
              select(Farm_ID, 
                     Reg, 
                     Color,
                     ends_with("2017")), 
            by = c("dam_asa_number" = "Reg")) %>% 
  filter(!is.na(CalvingSeason2017),
         !is.na(HairScore2017), 
         !is.na(calf_sex)) %>% 
  mutate(year = "2017") %>% 
  rename(Age = Age2017, CalvingSeason = CalvingSeason2017, ToxicFescue = ToxicFescue2017, HairScore = HairScore2017) %>% 
  select(-Shearing2017, - Comment2017, -Sold2017)
```

```{r}
sim_dam_long <- bind_rows(sim_dam_16, sim_dam_17)
```

 
## Using phenotypes only

```{r}
ww_ajdust_sim <- lm(adj_ww~1 + 
                  #as.numeric(HairScore) +
                  as.factor(Farm_ID) +
                  as.factor(calf_sex) +
                  as.factor(CalvingSeason) +
                    as.factor(year),
   data = sim_dam_long)

summary(ww_ajdust_sim)

#summary(ww_ajdust)

bound_residuals_sim <- bind_cols(sim_dam_long, 
                             as.data.frame(ww_ajdust_sim$residuals)) %>% 
  rename(calf_residual = `ww_ajdust_sim$residuals`)


#fill = c("#97B8C2")
bound_residuals_sim %>% 
  ggplot(aes(x = as.factor(HairScore), y = calf_residual)) +
  geom_boxplot(fill = c("#fde725ff"), alpha = 0.75) +
  labs(x = "Hair score", 
       y = str_wrap("205d + sex + farm + calving season + year adjusted weaning weight", 
                    width = 30),
       title = str_wrap("Registered Simmental 2016-2017 (n = 1704)", 
                        width = 45)
       )

ggsave("180917.sim_ww_boxplot.png", width = 10.5, height = 6.5, units = "in", dpi = 300)
```

```{r}

model.ww_hs <- summary(lm(adj_ww~1 + 
       as.numeric(HairScore2017) +
       as.factor(Farm_ID) +
       as.factor(calf_sex) +
       as.factor(CalvingSeason2017),
     data = sim_dam))

model.ww_color <- lm(adj_ww~1 + 
       as.factor(Color) +
       as.factor(Farm_ID) +
       as.factor(calf_sex) +
       as.factor(CalvingSeason2017),
     data = sim_dam)


anova(model.ww_color, model.ww_hs)

model_compare$`Pr(>F)`

AIC(model.ww_color, model.ww_hs)
```



## Using breeding values

> Pull in `sommer` generated breeding values

* Call to generate breeding values: 

`$fixed
HairScore ~ 1 + DateDeviation + CalvingSeason + Sex + Age + Farm_ID + 
    yr

$random
~g(international_id) + international_id
<environment: 0x14b61f38>

$rcov
~units `

```{r}
pre_adjust_multi <- readRDS("../../hair_gwas/pre_adjust/180509.pre_adjust_multi.RDS")
pre_adjust_multi$call
```

```{r}
#Pull out BLUPs
breeding_values <- pre_adjust_multi$u.hat[1] %>% 
  purrr::reduce(as.data.frame) %>% 
  #Take it from a matrix to a da
  as.data.frame() %>% 
  #make a column of international_id's (currently rownames)
  mutate(international_id = rownames(.)) %>% 
  rename(breeding_value = HairScore) 
```

> Pull in ASA calf phenotypes

```{r, warning = FALSE}
sim_legacy <- read_excel("data/180706.sim.Dam_Progeny_Report.xlsx") %>% 
  janitor::clean_names() %>% 
  #Change DOB and wean date to dates
  mutate_at(c("calf_dob", "wean_date"), as.Date) %>% 
  #Join progeny report to contemporary grouping data
  left_join(read_excel("data/180706.sim.Anm_CG_Data.xlsx") %>%
              janitor::clean_names() %>% 
              rename(calf_sex = sex, 
                     calf_asa_number = anm_nbr) %>%
              #Change sex codes
              mutate(calf_sex = replace(calf_sex, calf_sex == "C", "F"),
                     calf_sex = replace(calf_sex, calf_sex == "B" | calf_sex == "S", "M")) %>% 
              select(calf_sex, calf_asa_number)) %>% 
  #Join to phenotype data
  left_join(master %>% 
              select(Farm_ID, 
                     Reg, 
                     international_id), 
            by = c("dam_asa_number" = "Reg")) %>% 
  filter(international_id %in% breeding_values$international_id) %>% 
  #Join to breeding value data
  left_join(breeding_values, by = c("international_id")) %>% 
  filter(!is.na(breeding_value)) %>% 
  #Create a column of the average ww for each dam
  group_by(dam_asa_number) %>% 
  mutate(dam_avg_adj_ww = mean(adj_ww)) %>% 
  #Create a column of calf birth year
  mutate(birth_yr = str_extract(calf_dob, "^[:digit:]{4}"))  %>% 
  filter(!is.na(breeding_value) & !is.na(Farm_ID) & !is.na(calf_sex) & !is.na(birth_yr)) %>% 
  filter(adj_ww != 0)
            
```

```{r}

#Linear model predicting adjusted weaning weight from dam breeding value, birth year, farm, and sex
ww_ajdust_bv <- lm(adj_ww~1 +
                     as.factor(Farm_ID) +
                     as.factor(calf_sex) +
                     as.factor(birth_yr),
   data = sim_legacy)

summary(ww_ajdust_bv)

#Bind calf residuals to other data
bound_residuals_bv <- bind_cols(sim_legacy, 
                             as.data.frame(ww_ajdust_bv$residuals)) %>% 
  rename(calf_residual = `ww_ajdust_bv$residuals`)

bound_residuals_bv %>% 
  ggplot(aes(x = breeding_value, y = calf_residual)) +
  geom_point(color = c("#97B8C2")) +
  geom_smooth()
```


```{r}
#Regressing adjusted weaning weight on dam's hair shedding score breeding value
summary(
  lm(calf_residual~1 +
       as.numeric(breeding_value),
   data = bound_residuals_bv)
  )



#What's up with chunk of observations below 400?
bound_residuals_bv %>% 
  filter(calf_residual < -350)

```


# New distribution map:

```{r}
#Base USA map
usa <- borders("state", regions = ".", fill = "white", colour = "black")
```

```{r}
master %>% 
         filter(!is.na(Breed), Breed != "CHIA", Breed != "MAAN", Breed != "BRN") %>% 
  filter(!is.na(HairScore2016) | !is.na(HairScore2017)) %>% 
ggplot(aes(x = LNG,
           y= LAT,
           colour = factor(Breed))) +
  usa +
  geom_count(alpha = 0.7, 
             #Shift position of overlapping points so that they can be seen better
             position = position_jitterdodge(dodge.width = 0.8, 
                                             jitter.width = 0.5, 
                                             jitter.height = -0.4)) +
  scale_size_area(max_size = 8, 
                  #Take off x & y axis border lines
                  guide = 'none') +
  scale_color_viridis(discrete = TRUE,
                      labels = c("Angus: 2,930", "Red Angus: 693", "Charolais: 285", "Cross: 426", "Gelbvieh: 280", "Hereford: 1,273", "Shorthorn: 276", "Simmental: 1,827")) +
  #Make the key have 4 rows and 2 columns
  guides(col = guide_legend(nrow = 4, 
               ncol = 2,
               title = "Breed")) +
  #coord_fixed() +
  theme_map() +
  #Set the "anchoring point" of the legend (bottom-left is 0,0; top-right is 1,1)
  #Put bottom-left corner of legend box in bottom-left corner of graph
  theme(legend.justification=c(0,.05), 
        legend.position=c(0,0), 
        legend.key.size = unit(1, "cm")) 


ggsave("180713.breed_map.png", width = 10, height = 6.5, units = "in", dpi = 300)
```
