---
title: "Thompson field day 2018"
author: "Harly Durbin"
date: "9/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(stringi)
library(GGally)
library(lubridate)
library(cowplot)
library(viridis)
```

```{r}
letter_yr <- tribble(~letter, ~Age2018,
                     "A", 5,
       "B", 4,
       "C", 3,
       "D", 2,
       "E", 1,
       "M", 16,
       "N", 15,
       "P", 14,
       "R", 13,
       "S", 12,
       "T", 11,
       "U", 10,
       "W", 9,
       "X", 8,
       "Y", 7,
       "Z", 6)
```


# Data formatting

## Thompson

### Hair shedding metadata: wide format

```{r}

#Create a list of file paths for Thompson data
filepaths <-
  combine(
    list.files(
      path = "~/Box Sync/HairShedding/ReportedData/2016",
      pattern = "UMCT",
      full.names = TRUE
    ),
    list.files(
      path = "~/Box Sync/HairShedding/ReportedData/2017",
      pattern = "UMCT",
      full.names = TRUE
    ),
    list.files(
      path = "~/Box Sync/HairShedding/ReportedData/2018",
      pattern = "UMCT",
      full.names = TRUE
    )
  )


data_recording_wide <- filepaths %>%
  #Name the elements of the list based on a stripped down version of the filepath
  set_names(nm = (basename(.) %>%
                    tools::file_path_sans_ext())) %>%
  #Create a list of data frames using readxl
  map(read_excel, sheet = 1) %>%
  #Select the columns I need
  map(select, Sex, Color, Animal_ID, DateScoreRecorded:ToxicFescue) %>%
  map(
    mutate,
    Color = "Black",
    CalvingSeason = "Spring",
    ToxicFescue = "Yes"
  ) %>%
  #Keep only observations with hair scores, remove bulls
  map(filter, !is.na(HairScore), Sex == "F") %>%
  #imap uses the data frame's name in the list as y
  #Mutate each data frame in the list to append the year to the end of a few columns based on name
  #https://stackoverflow.com/questions/29948876/adding-prefix-or-suffix-to-most-data-frame-variable-names-in-piped-r-workflow
  imap(~rename_at(.x,
                    vars(
                      one_of("DateScoreRecorded", "HairScore", "Age")
                    ),
                    function(i)
                      str_c(i, str_extract(
                        .y, "(?<=_)[[0-9]]+"
                      )))) %>%
  #Collapse the dataframes to one dataframe
  reduce(full_join) %>%
  #Remove Age2018 column
  select(-Age2018) %>%
  #I can't remember why I did this oops. To determine unknown ages based on the letter in their ID?
  mutate(letter = str_extract(Animal_ID, "(?<=1)[A-Z]")) %>%
  left_join(letter_yr, by = c("letter")) %>%
  select(-letter) %>%
  mutate(Age2017 = if_else(
    is.na(Age2017) & !is.na(Age2018),
    Age2018 - 1,
    as.double(Age2017)
  )) %>%
  mutate(Age2016 = if_else(
    is.na(Age2016) & !is.na(Age2017),
    Age2017 - 1,
    as.double(Age2016)
  ))


```


### Hair shedding metadata: long format (one observation per year per row)


```{r}

filepaths <-
  combine(
    list.files(
      path = "~/Box Sync/HairShedding/ReportedData/2016",
      pattern = "UMCT",
      full.names = TRUE
    ),
    list.files(
      path = "~/Box Sync/HairShedding/ReportedData/2017",
      pattern = "UMCT",
      full.names = TRUE
    ),
    list.files(
      path = "~/Box Sync/HairShedding/ReportedData/2018",
      pattern = "UMCT",
      full.names = TRUE
    )
  )


data_recording_long <- filepaths %>%
  #Name the elements of the list based on a stripped down version of the filepath
  set_names(nm = (basename(.) %>%
                    tools::file_path_sans_ext())) %>%
  #Create a list of data frames using readxl
  map(read_excel, sheet = 1) %>%
  map(select, Sex, Color, Animal_ID, DateScoreRecorded:ToxicFescue) %>%
  map(
    mutate,
    Color = "Black",
    CalvingSeason = "Spring",
    ToxicFescue = "Yes",
    letter = str_extract(Animal_ID, "(?<=1)[A-Z]")
  ) %>%
  map(filter,!is.na(HairScore), Sex == "F") %>%
  #imap uses the data frame's name in the list as y
  imap( ~ mutate(.x, year = str_extract(.y, "(?<=_)[[0-9]]+"))) %>%
  reduce(bind_rows) %>%
  left_join(letter_yr, by = c("letter")) %>%
  mutate(Age = if_else(year == "2016",
                       Age2018 - 2,
                       Age)) %>%
  mutate(Age = if_else(year == "2017",
                       Age2018 - 1,
                       Age)) %>%
  mutate(Age = if_else(year == "2018",
                       Age2018,
                       Age)) %>%
  select(-Age2018,-letter)


```

### Pedigree & weaning weight data

```{r}

#Get pedigree information including BIR numbers (to be used as Reg numbers)
#list of dataframes; a dataframe for each file
pedigree <-
  list.files(
    "~/googledrive/research_extension/hair_shed/hair_shed/hair_project/data",
    pattern = "TRC_Pedigree",
    full.names = TRUE
  ) %>%
  #Name the elements of the list based on a stripped down version of the filepath
  set_names(nm = (basename(.) %>%
                    tools::file_path_sans_ext())) %>%
  #Create a list of data frames using readxl
  map_df(read_excel, sheet = 1, trim_ws = TRUE) %>%
  janitor::clean_names() %>%
  select(anm_tag:reg_no, sex, sire_reg_no, dam_tag, dam_reg_no) %>%
  #Formate birth date as a date
  mutate(anm_birth_date = mdy(anm_birth_date)) %>%
  #Pull in historical weaning weight/growth phenotype data
  left_join(
    list.files(
      "~/Box Sync/GeneMax/RawData/",
      pattern = "Growth.*.xlsx",
      full.names = TRUE
    ) %>%
      set_names(nm = (
        basename(.) %>%
          tools::file_path_sans_ext()
      )) %>%
      #Create a list of data frames using readxl
      map_df(read_excel, sheet = 1, trim_ws = TRUE) %>%
      janitor::clean_names() %>%
      select(anm_tag, anm_birth_date, reg_no, sex, birth_wt:yr_ratio) %>%
      mutate(anm_birth_date = mdy(anm_birth_date))
  ) %>%
  rename_at(vars(-starts_with("dam")), ~ str_c("calf", "_", .)) %>%
  #Format sex codes
  mutate(
    calf_sex = replace(calf_sex, calf_sex == "C", "F"),
    calf_sex = replace(calf_sex, calf_sex == "B" |
                         calf_sex == "S", "M")
  )
    
    

```


```{r}

growth <- bind_rows(
  left_join(
    data_recording_long %>%
      filter(year == "2016"),
    pedigree %>%
      filter(year(calf_anm_birth_date) == 2016),
    by = c("Animal_ID" = "dam_tag")
  ),
  left_join(
    data_recording_long %>%
      filter(year == "2017"),
    pedigree %>%
      filter(year(calf_anm_birth_date) == 2017),
    by = c("Animal_ID" = "dam_tag")
  ),
  left_join(
    data_recording_long %>%
      filter(year == "2018"),
    pedigree %>%
      filter(year(calf_anm_birth_date) == 2018),
    by = c("Animal_ID" = "dam_tag")
  )
)
```


## Red Angus

### 2016

```{r}
anr_ww_16 <- read_csv("data/180503_anr_ww.csv") %>%
  mutate(
    regisno = as.character(regisno),
    dam_reg = as.character(dam_reg),
    birth_date = mdy(birth_date),
    wean_date = mdy(wean_date),
    dam_dob = mdy(dam_dob),
    year_date = mdy(year_date)
  ) %>%
  dplyr::select(
    regisno,
    prefix,
    sex,
    birth_date,
    ownr,
    breeder,
    sire_reg,
    dam_reg,
    dam_dob,
    birth_wt,
    birth_adj,
    wean_date,
    wean_wt,
    wean_adj,
    year_date,
    year_wt,
    year_adj
  ) %>%
  left_join(master %>%
              select(Farm_ID,
                     Reg,
                     Color,
                     ends_with("2016")),
            by = c("dam_reg" = "Reg")) %>%
  filter(year(wean_date) == "2016") %>%
  filter(!is.na(wean_adj)) %>%
  rename(calf_sex = sex) %>%
  mutate(
    calf_sex = replace(calf_sex, calf_sex == "C", "F"),
    calf_sex = replace(calf_sex, calf_sex == "B" |
                         calf_sex == "S", "M")
  ) %>%
  mutate(year = "2016") %>%
  rename(
    Age = Age2016,
    CalvingSeason = CalvingSeason2016,
    ToxicFescue = ToxicFescue2016,
    HairScore = HairScore2016
  ) %>%
  select(-Shearing2016,-Comment2016)
```

### 2017

```{r}
anr_ww_17 <- read_csv("data/180503_anr_ww.csv") %>%
  mutate(
    regisno = as.character(regisno),
    dam_reg = as.character(dam_reg),
    birth_date = mdy(birth_date),
    wean_date = mdy(wean_date),
    dam_dob = mdy(dam_dob),
    year_date = mdy(year_date)
  ) %>%
  dplyr::select(
    regisno,
    prefix,
    sex,
    birth_date,
    ownr,
    breeder,
    sire_reg,
    dam_reg,
    dam_dob,
    birth_wt,
    birth_adj,
    wean_date,
    wean_wt,
    wean_adj,
    year_date,
    year_wt,
    year_adj
  ) %>%
  left_join(master %>%
              select(Farm_ID,
                     Reg,
                     Color,
                     ends_with("2017")),
            by = c("dam_reg" = "Reg")) %>%
  filter(year(wean_date) == "2017") %>%
  filter(!is.na(wean_adj)) %>%
  rename(calf_sex = sex) %>%
  mutate(
    calf_sex = replace(calf_sex, calf_sex == "C", "F"),
    calf_sex = replace(calf_sex, calf_sex == "B" |
                         calf_sex == "S", "M")
  ) %>%
  mutate(year = "2017") %>%
  rename(
    Age = Age2017,
    CalvingSeason = CalvingSeason2017,
    ToxicFescue = ToxicFescue2017,
    HairScore = HairScore2017
  ) %>%
  select(-Shearing2017,-Comment2017,-Sold2017)
```

### Combined

```{r}
anr_long <- bind_rows(anr_ww_16, anr_ww_17) %>%
  filter(
    !is.na(Farm_ID) &
      !is.na(calf_sex) &
      !is.na(CalvingSeason) &
      !is.na(wean_adj) &
      !is.na(HairScore)
  )
```



## National-level Angus

```{r}

angus_adjust <- bind_rows(
  master %>%
    filter(
      Breed == "AN" &
        !is.na(HairScore2017) &
        !is.na(CalvingSeason2017) &
        !is.na(DateScoreRecorded2017)
    ) %>%
    mutate(
      Group = "National",
      DateDeviation2016 =  as.Date(DateScoreRecorded2016) - ymd("2016-05-01"),
      DateDeviation2017 =  as.Date(DateScoreRecorded2017) - ymd("2017-05-01"),
      HairScore2017 = as.numeric(HairScore2017),
      HairScore2016 = as.numeric(HairScore2016)
    ),
  data_recording_wide %>%
    mutate(
      Group = "Thompson",
      Farm_ID = "TRC",
      DateDeviation2016 =  as.Date(DateScoreRecorded2016) - ymd("2016-05-01"),
      DateDeviation2017 =  as.Date(DateScoreRecorded2017) - ymd("2017-05-01"),
      CalvingSeason2017 = "Spring"
    ) %>%
    filter(
      !is.na(HairScore2017) &
        !is.na(CalvingSeason2017) &
        !is.na(DateDeviation2017)
    )
)

```


# Explore

## Thompson

```{r}
growth %>% 
  group_by(HairScore) %>% 
  tally()

growth %>% 
  ggplot(aes(x = HairScore)) + 
  geom_histogram() +
  scale_fill_viridis(discrete = TRUE)
```

## `sommer` generated breeding values

```{r}

pre_adjust_multi <- readRDS("pre_adjust_multi.RDS")

```


```{r}

pre_adjust_multi$data %>%
  group_by(Farm_ID) %>%
  tally()

```

```{r}

u.hat_multi <- pre_adjust_multi$u.hat[1] %>%
  purrr::reduce(as.data.frame) %>%
  as.data.frame()

u.hat_multi[, 2] <- row.names(u.hat_multi)

u.hat_multi <- u.hat_multi %>%
  rename(u.hat = HairScore, international_id = V2)

```

```{r}
suma <- summary(pre_adjust_multi)$var.comp.table

sum(suma[1,1]/sum(suma[,1]))
```

```{r}

cov_resid <- left_join(cov, u.hat_multi, by = c("international_id"))

```


# Model/pre-adjust

## Thompson

```{r}
grow2 <- growth %>% 
  filter(
           !is.na(Age) &
           !is.na(HairScore) &
           !is.na(calf_sex) &
           !is.na(year) & 
           !is.na(calf_wn_adj_wt))
  

thompson_preadjust <- summary(lm(calf_wn_adj_wt~1 +
             Age +
             calf_sex +
             year, 
   data = grow2))

```



## Red Angus


```{r}

ww_ajdust_anr <- lm(
  wean_adj ~ 1 +
    #as.numeric(HairScore) +
    as.factor(Farm_ID) +
    as.factor(calf_sex) +
    as.factor(CalvingSeason) +
    as.factor(year),
  data = anr_long
)

summary(ww_ajdust_anr)

```


## National-level Angus

```{r}
angus_adjust_lm <- lm(
  as.numeric(HairScore2017) ~ 1 +
    as.factor(CalvingSeason2017) +
    as.numeric(DateDeviation2017) +
    as.factor(Farm_ID),
  data = angus_adjust
)

length(angus_adjust_lm$residuals)

```

# Plot

## Thompson specific

### Thompson adjusted weaning weight vs. dam hair shedding score

```{r}
library(colormap)
colormap(colormap = colormaps$viridis, nshades = 8)

thompson_residuals <-
  bind_cols(grow2,                        as.data.frame(thompson_preadjust$residuals)) %>%
  rename(calf_residual = `thompson_preadjust$residuals`)

thompson_residuals %>%
  ggplot(aes(x = as.factor(HairScore), y = calf_residual)) +
  geom_boxplot(fill = c("#440154ff"), alpha = 0.75) +
  labs(
    x = "Hair score",
    y = str_wrap("205d + sex + year adjusted weaning weight", width = 30),
    title = "Thompson 2016-2017 (n = 312)"
  )


ggsave(
  "180917.thompson_ww.png",
  width = 10.5,
  height = 6.5,
  units = "in",
  dpi = 300
)
```


## Red Angus

```{r}


bound_residuals_anr <- bind_cols(anr_long,
                                 as.data.frame(ww_ajdust_anr$residuals)) %>%
  rename(calf_residual = `ww_ajdust_anr$residuals`)


#fill = c("#97B8C2")
bound_residuals_anr %>%
  filter(!is.na(HairScore)) %>%
  ggplot(aes(x = as.factor(HairScore), y = calf_residual)) +
  geom_boxplot(fill = c("#46307cff"), alpha = 0.75) +
  labs(
    x = "Hair score",
    y = str_wrap(
      "205d + sex + farm + calving season + year adjusted weaning weight",
      width = 30
    ),
    title = str_wrap("Registered Red Angus 2016-2017 (n = 241)",
                     width = 45)
  )

ggsave(
  "180917.anr_ww_boxplot.png",
  width = 10.5,
  height = 6.5,
  units = "in",
  dpi = 300
)

```

## Thompson hair shedding scores vs. national scale Angus hair shedding scores


```{r}
angus_residuals <- bind_cols(angus_adjust,
                             as.data.frame(angus_adjust_lm$residuals)) %>%
  rename(hair_residual = `angus_adjust_lm$residuals`)

ggplot(data = angus_residuals, aes(x = as.numeric(hair_residual), fill = Group)) +
  geom_density(alpha = 0.25) +
  scale_fill_viridis(discrete = TRUE) +
  labs(
    title = str_wrap("Thompson Farm vs. national Angus hair scores", width = 30),
    y = "Density",
    x = str_wrap(
      "Calving season + date scored + farm adjusted hair shedding score",
      width = 45
    )
  )

ggsave(
  "180916.thompson_density.png",
  width = 10.5,
  height = 6.5,
  units = "in",
  dpi = 300
)
```


## Other

### Hair shedding score by breed

```{r}

master %>%
  filter(!is.na(Breed), Breed != "CROS") %>%
  ggplot(aes(x = as.numeric(HairScore2017), fill = Breed)) +
  geom_density(alpha = 0.25) +
  scale_fill_viridis(discrete = TRUE) +
  labs(
    title = str_wrap("Hair shedding score by breed", width = 30),
    y = "Density",
    x = str_wrap(
      "Calving season + date scored + farm adjusted hair shedding score",
      width = 45
    )
  )

```

### Hair shedding score breeding values from `sommer` by breed

## Denstiy

```{r}

as_data_frame(pre_adjust_multi$data) %>%
  mutate(international_id = as.character(international_id)) %>%
  left_join(master %>%
              select(international_id, Breed)) %>%
  filter(!is.na(Breed) &
           Breed != "CHIA" & Breed != "CROS" & Breed != "MAAN") %>%
  left_join(as_data_frame(u.hat_multi)) %>%
  ggplot(aes(x = as.numeric(u.hat), fill = Breed)) +
  geom_density(alpha = 0.25) +
  scale_fill_viridis(discrete = TRUE) +
  labs(
    title = str_wrap("Hair shedding score EPD by breed", width = 30),
    y = "Density",
    x = str_wrap("Hair shedding score EPD", width = 45)
  )

ggsave(
  "180916.thompson_epd_by_breed.png",
  width = 10.5,
  height = 6.5,
  units = "in",
  dpi = 300
)


```


## Boxplot

```{r}

as_data_frame(pre_adjust_multi$data) %>%
  mutate(international_id = as.character(international_id)) %>%
  left_join(master %>%
              select(international_id, Breed)) %>%
  filter(!is.na(Breed) &
           Breed != "CHIA" & Breed != "CROS" & Breed != "MAAN") %>%
  left_join(as_data_frame(u.hat_multi)) %>% 
  ggplot(data = test, aes(x = as.numeric(u.hat), fill = Breed)) +
  geom_density(alpha = 0.25) +
  scale_fill_viridis(discrete = TRUE) +
  labs(
    title = str_wrap("Hair shedding score EPD by breed", width = 30),
    y = "Density",
    x = str_wrap("Hair shedding score EPD", width = 45)
  )

ggsave(
  "180916.thompson_epd_by_breed.png",
  width = 10.5,
  height = 6.5,
  units = "in",
  dpi = 300
)

```

